---
title: "Data preparation"
author: "Willem"
date: "2025-06-06"
output: html_document
---
# PLANT-POL
# 1 START
Load
```{r}
library(tidyverse)
library(rgbif)
library(kewr)
library(ggpubr)
library(sf)
library(maps)
library(ggnewscale)
library(bipartite)
library(ggbipart)
library(networkD3)
```


```{r}
polplant <- read.csv2("../Data/2024-145_BE_metWP_sws+metToestemming.csv")
```


# 2 Data cleaning
## 2.1 Plants
Names
```{r}
# get all names
names <- polplant |> group_by(waardplant) |> summarise(n=n())
names$original <- names$waardplant

# remove "spec"
names$waardplant <- gsub("\\b(spec\\.)\\b", "", names$waardplant)
names$waardplant <- trimws(names$waardplant)
indet <- names$waardplant[grepl(" indet.", names$waardplant)]
names$waardplant <- gsub(" indet.","", names$waardplant)

## PLANTS
# check against gbif for kingdom
gbif <- name_backbone_checklist(names$waardplant)

# this sometimes gives errors, here is patchwork option
# also not perfect --> no idea why
gbif <- name_backbone_checklist(names$waardplant[1])
n <- nrow(names)
n <- ceiling(n/20)
for( i in 18: n){
  begin <- 2 + (i-1)*20
  end <- 1+ i *20
  data <- name_backbone_checklist(names$waardplant[begin:end])
  gbif <- bind_rows(gbif, data)
  Sys.sleep(20)
}

# check which are not plant in gbif
non_plants <- gbif[!gbif$kingdom =="Plantae",]
non_plants <- non_plants[!is.na(non_plants$verbatim_name),]

# check which are not plant in powo
non_plants_powo <- match_knms(non_plants$verbatim_name)
non_plants_powo <- non_plants_powo$results
non_plants_powo_plants <- non_plants[lengths(non_plants_powo)>2,]
non_plants_powo_plants <- non_plants_powo_plants[! non_plants_powo_plants$verbatim_name %in% c("Anthophila indet.", "Osmia (Osmia)","Anthophila"),]
non_plants_powo <- non_plants[! non_plants$verbatim_name %in% non_plants_powo_plants$verbatim_name,]

# remove non-plants from name list
names <- names[!names$waardplant %in% non_plants_powo$verbatim_name,]

# check in powo 
names_powo <- match_knms(names$waardplant)
names_powo <- names_powo$results


# safe depending on types
safe_names <- c()
multiple <- c()
none <- c()
# results can be false true multiple_matches or ""
# all together also 1697
for(i in 1: length(names_powo)){
  if(names_powo[[i]][[2]]== "true"){
    safe_names <- rbind(safe_names, names_powo[[i]][[1]])
  } else if(names_powo[[i]][[2]]=="multiple_matches"){
    multiple <- rbind(multiple, names_powo[[i]][[1]])
  } else if(names_powo[[i]][[2]]=="false"){
    none <- rbind(none, names_powo[[i]][[1]])
  }
}

i=1

################
# SAFE
safe_powo <- list()
saved=T
if(!saved){
for(i in 1: length(safe_names)){
    # powo
    r<- search_powo(safe_names[i])
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      safe_powo[[i]] <- tmp
    } else {
      safe_powo[[i]] <- NA
    }
    #write_rds(safe_powo, "../Extra/Names/powo_synonyms.R")
}
  } else {
  safe_powo <- read_rds("../Extra/Names/powo_synonyms.R")
}
i=11

not_safe_names <- safe_names[!lengths(safe_powo) > 2]
not_saf_names_powo <- safe_powo[!lengths(safe_powo) > 2]
safe_names <- safe_names[!lengths(safe_powo) <2]
safe_powo <- safe_powo[!lengths(safe_powo) < 2]


# get the correct and accepted names
synonym_safe <- data.frame(old = character(), new= character(), accepted= character())
for(i in 1: length(safe_powo)){
  dat <- safe_powo[[i]]
  acc_name <- NA
  acc <- NA
  old <- NA
  if(nrow(dat) == 1){
    synonym_safe <- rbind(synonym_safe, data.frame(old= safe_names[i], new = dat$name, accepted= dat$accepted))
  } else {
    acc_name <- dat$name[dat$accepted==T]
    acc_name <- ifelse(any(acc_name %in% safe_names[i]), acc_name[acc_name %in% safe_names[i]], acc_name)
    acc <- dat$accepted[dat$accepted==T]
    acc <- ifelse(length(acc)>1, any(dat$accepted[dat$name %in% acc_name]), acc)
    old <- safe_names[i]
    synonym_safe <- rbind(synonym_safe, data.frame(old= old, new = acc_name, accepted= acc))
  }
}


# give powo names also to name list
names$powo <- synonym_safe$new[match(names$waardplant, synonym_safe$old)]
names$genus <- gbif$genus[match(names$powo, gbif$canonicalName)]
names$family <- gbif$family[match(names$powo, gbif$canonicalName)]


################
# NONE
# search powo and store results in list if we find a direct match
none_powo <- list()
saved = T
if(!saved){
for(i in 1: length(none)){
    # powo
    r<- search_powo(none[i])
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      none_powo[[i]] <- tmp
    } else {
      none_powo[[i]] <- NA
    }
    #write_rds(none_powo, "../Extra/Names/powo_no_link.R")
}
} else {
  none_powo <- read_rds("../Extra/Names/powo_no_link.R")
}

# only get all with data
none_name <- none[!lengths(none_powo)<2]
none_powo_all <- none_powo
none_powo <- none_powo[!lengths(none_powo)<2]

# get the correct and accepted names
none_safe <- data.frame(old = character(), new= character(), accepted= character())
for(i in 1: length(none_powo)){
  dat <- none_powo[[i]]
  acc_name <- NA
  acc <- NA
  old <- NA
  if(nrow(dat) == 1){
    none_safe <- rbind(none_safe, data.frame(old= none_name[i], new = dat$name, accepted= dat$accepted))
  } else {
    acc_name <- dat$name[dat$accepted==T]
    acc <- dat$accepted[dat$accepted==T]
    old <- none_name[i]
    none_safe <- rbind(none_safe, data.frame(old= old, new = acc_name, accepted= acc))
  }
}
none_safe[none_safe$old %in% c("Astereae","Eupatorieae","Heliantheae"), 2:3] <- NA

dup <- none_safe[duplicated(none_safe$old),]
gbif_dup <- name_backbone_checklist(dup$old)
any(gbif_dup$rank != "FAMILY" & gbif_dup$rank != "KINGDOM")

none_safe$new[none_safe$old %in% dup$old] <- none_safe$old[none_safe$old %in% dup$old]

names$powo[is.na(names$powo)] <- none_safe$new[match(names$waardplant[is.na(names$powo)], none_safe$old)]

################
# MULTIPLE
# search powo and store results in list if we find a direct match
multiple_powo <- list()
if(!saved){
for(i in 1: length(multiple)){
    # powo
    r<- search_powo(multiple[i])
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      multiple_powo[[i]] <- tmp
    } else {
      multiple_powo[[i]] <- NA
    }
    write_rds(multiple_powo, "../Extra/Names/powo_multiple.R")
}
  } else {
  multiple_powo <- read_rds("../Extra/Names/powo_multiple.R")
}

# only get all with data
multiple_name <- multiple[!lengths(multiple_powo)<2]
multiple_powo_all <- multiple_powo
multiple_powo <- multiple_powo[!lengths(multiple_powo)<2]

# get the correct and accepted names
multiple_safe <- data.frame(old = character(), new= character(), accepted= character())
for(i in 1: length(multiple_powo)){
  dat <- multiple_powo[[i]]
  acc_name <- NA
  acc <- NA
  old <- NA
  if(nrow(dat) == 1){
    multiple_safe <- rbind(multiple_safe, data.frame(old= multiple_name[i], new = dat$name, accepted= dat$accepted))
  } else {
    acc_name <- dat$name[dat$accepted==T]
    acc_name <- ifelse(any(acc_name %in% multiple_name[i]), acc_name[match(acc_name, multiple_name[i])], acc_name)
    acc <- dat$accepted[dat$accepted==T]
    acc <- ifelse(length(acc)>1, any(dat$accepted[dat$name %in% acc_name]), acc)
    old <- multiple_name[i]
    multiple_safe <- rbind(multiple_safe, data.frame(old= old, new = acc_name, accepted= acc))
  }
}

names$powo[is.na(names$powo)] <- multiple_safe$new[match(names$waardplant[is.na(names$powo)], multiple_safe$old)]

# check which not yet
multiple_not <- multiple_safe[is.na(multiple_safe$new),]
# adapt
multiple_not$new <- c("Arum","Camellia","Centaurea nigra","Cirsium palustre","Erodium","Forsythia","Geranium","Hypericum hirsutum","Ilex","Jacobaea vulgaris", 
                      "Linum usitatissimum","Spiraea japonica")
# safe
names$powo[is.na(names$powo)] <- multiple_not$new[match(names$waardplant[is.na(names$powo)], multiple_not$old)]
```

Check not safe names
```{r}
not_safe_gbif <- name_backbone_checklist(not_safe_names)

multiple_species <- not_safe_gbif[grepl("\\+", not_safe_gbif$verbatim_name),]

# check with gbif
not_safe_gbif <- name_backbone_checklist(not_safe_names)
not_safe_gbif_acc <- not_safe_gbif[!is.na(not_safe_gbif$status),]

# give accepted names
names$powo[names$waardplant %in% not_safe_gbif_acc$verbatim_name] <- 
  not_safe_gbif_acc$scientificName[match(names$waardplant[names$waardplant %in% not_safe_gbif_acc$verbatim_name], not_safe_gbif_acc$verbatim_name)]


not_safe_gbif <- not_safe_gbif[is.na(not_safe_gbif$status),]
```



Check remaining names
```{r}
# check and remove some names incorrectly added
lacking <- names[is.na(names$powo),]

lacking$powo <- c("Aconitum × cammarum","Aesculus x carnea","Allium christophii","Allium vineale","Eriocapitella × hybrida","Anthemideae",
                  "Arctium minus subsp. minus","Arctium x ambiguum","Astereae","Berberis × ottawensis","Brassica napus subsp. napus","Buddleja x weyeriana",
                  "Calamintha nepeta subsp. nepeta","Cardueae","Caryopteris x clandonensis","Centaurea cyanus","Cichorieae",
                  "Citrus x aurantium","Clinopodium nepeta subsp. spruneri","Crocus x helenae","Crocus x luteus","Echinops banaticus","Erica x darleyensis",
                  "Erodium cicutarium subsp. cicutarium","Eupatorieae","Fallopia x bohemica","Ficaria verna","Fragaria x ananassa","Gaillardia x grandiflora",
                  "Geranium x cantabrigiense","Geranium x magnificum","Geranium x oxonianum","Heliantheae","Helianthus x laetiflorus",
                  "Heracleum sphondylium subsp. sphondylium","Hyacinthoides x massartiana","Hyacinthoides x massartiana","Hypericum x desetangsii",NA,
                  "Knautia arvensis","Lonicera fragrantissima subsp. fragrantissima","Medicago x varia","Mentha x verticillata",
                  "Mentha x piperita","Mentha x rotundifolia","Micrasterias decemdentata",NA,NA, "Narcissus pseudonarcissus subsp. pseudonarcissus",
                  "Nepeta x faassenii","Ononis spinosa subsp. procurrens","Osmanthus x burkwoodii","Petunia × atkinsiana",NA, "Platanus x hispanica",
                  "Populus x canadensis","Poterium sanguisorba subsp. sanguisorba","Prunus avium subsp. avium","Prunus domestica subsp. domestica",
                  "Prunus x fruticans","Rhododendron x praecox","Ribes x nidigrolaria","Rosa subsect. Caninae","Rosa subsect. Rubigineae",
                  "Rubus sect. Corylifolii","Rubus sect. Rubus","Rubus sect. Rubus","Salix","Salix alba","Salix babylonica","Salix cinerea subsp. cinerea",
                  "Salix hybrid s.l.","Salix x ambigua","Salix x friesiana","Salix x guinieri","Salix x reichardtii","Salix x rubra",
                  "Salix × pendulina nothof. tristis","Salix gmelinii","Salix gmelinii","Salvia x sylvestris","Scilla sect. Chionodoxa",
                  "Sempervivum sect. Jovibarba","Solanum nigrum subsp. nigrum","Symphoricarpos x chenaultii","Symphyotrichum x versicolor",
                  "Symphytum grandiflorum", "Symphytum x hidcotense","Symphytum x uplandicum","Taraxacum officinale s.l.","Taraxacum sect. Ruderalia",
                  "Teucrium x lucidrys","Tilia x europaea","Glandularia × hybrida","Veronica chamaedrys var. chamaedrys","Vicia sativa subsp. angustifolia",
                  "Vicia tetrasperma subsp. tetrasperma" ,"Viola canina","Viola x wittrockiana","Vitis vinifera subsp. vinifera")

names$powo[is.na(names$powo)] <- lacking$powo
names <- names[!is.na(names$powo),]
names$powo[names$powo =="Brassica spec. + Sinapis"] <- "Brasica"
```


Again GBIF
```{r}
gbif_powo <- name_backbone_checklist(names$powo)

names$gbif <- gbif_powo$species[match(names$powo, gbif_powo$verbatim_name)]

# check differences
dif <- names[names$powo != names$gbif,]
dif <- dif[!is.na(dif$waardplant),]

# subspecies are currently merged, which is not really the purpose here
dif_subsp <- dif[dif$powo %in% gbif_powo$verbatim_name[gbif_powo$rank=="SUBSPECIES"],]
gbif_powo_subsp <- gbif_powo[gbif_powo$rank %in% c("SUBSPECIES"),]
# we will give them their cononical names
vegdata_name <- vegdata::parse.taxa(gbif_powo_subsp$scientificName)
vegdata_name[!vegdata_name$rank.suff=="subsp.",4:6] <- c("subsp.","australis", "Cornus sanguinea subsp. australis")
gbif_powo_subsp$vegdata <- vegdata_name$scientificName
# are there differences
gbif_powo_subsp[!gbif_powo_subsp$verbatim_name %in% gbif_powo_subsp$vegdata, ]

# assign names for subspecies again
dif$gbif[dif$powo %in% gbif_powo_subsp$verbatim_name] <- gbif_powo_subsp$vegdata

# assign to names vector
names$powo[names$waardplant %in% dif$waardplant] <- dif$gbif[match(names$waardplant[names$waardplant %in% dif$waardplant],dif$waardplant)]
names$gbif[names$waardplant %in% dif$waardplant] <- dif$gbif[match(names$waardplant[names$waardplant %in% dif$waardplant],dif$waardplant)]



gbif_again <- name_backbone_checklist(names$powo)

# look at synonyms
synonyms = gbif_again[gbif_again$status != "ACCEPTED",]
synonyms <- synonyms[!is.na(synonyms$usageKey),]
synonyms <- synonyms[!is.na(synonyms$acceptedUsageKey),]

synonyms_accepted <- name_usage(synonyms$acceptedUsageKey[1])$data

for( i in 2: nrow(synonyms)){
  tmp <- name_usage(synonyms$acceptedUsageKey[i])$data
  synonyms_accepted <- bind_rows(synonyms_accepted, tmp)
}

# get subspecies
synonyms_accepted_subsp <-  synonyms_accepted[ synonyms_accepted$rank %in% c("SUBSPECIES" , "VARIETY"),]
# we will adapt their cononical names
vegdata_name <- vegdata::parse.taxa(synonyms_accepted_subsp$scientificName)
synonyms_accepted$canonicalName[synonyms_accepted$rank %in% c("SUBSPECIES" , "VARIETY")] <- vegdata_name$scientificName
# adapt in overall matrix
synonyms$canonicalName <- synonyms_accepted$canonicalName

# assign to name matrix
names$gbif[names$powo %in% synonyms$verbatim_name] <- synonyms$canonicalName[match(names$powo[names$powo %in% synonyms$verbatim_name], synonyms$verbatim_name )]
names$powo[names$powo %in% synonyms$verbatim_name] <- synonyms$canonicalName[match(names$powo[names$powo %in% synonyms$verbatim_name],
                                                                                   synonyms$verbatim_name )]

# look whether all synonyms are changed
gbif_again <- name_backbone_checklist(names$powo)
synonyms = gbif_again[gbif_again$status != "ACCEPTED",]
synonyms <- synonyms[!is.na(synonyms$usageKey),]
# these are all doubtful

# check again for genera
names$genus[is.na(names$genus)] <- gbif_powo$genus[match(names$powo[is.na(names$genus)], gbif_powo$verbatim_name)]
names$family[is.na(names$family)] <- gbif_powo$family[match(names$powo[is.na(names$family)], gbif_powo$verbatim_name)]
# check again for genera
names$genus[is.na(names$genus)] <- gbif_again$genus[match(names$powo[is.na(names$genus)], gbif_again$verbatim_name)]
names$family[is.na(names$family)] <- gbif_again$family[match(names$powo[is.na(names$family)], gbif_again$verbatim_name)]

# check first with powo
genus_powo <- tidy(match_knms(names$powo[is.na(names$genus)]))
genus_powo <- genus_powo[!is.na(genus_powo$ipni_id),]
genus_powo$genus <- vegdata::parse.taxa(genus_powo$submitted)$genus

names$genus[is.na(names$genus)] <- genus_powo$submitted[match(names$powo[is.na(names$genus)], genus_powo$submitted)]

```


Genus
```{r}
# which have no genus or family
no_genus <- names[is.na(names$genus),]
no_family <- names[is.na(names$family),]

# first use gbif again
gbif <-  name_backbone_checklist(no_genus$powo)
# no longer relevant because previous move
#no_genus$genus <- gbif$genus
no_genus$family <- gbif$family

# assign
names[is.na(names$genus),] <- no_genus

no_family <- names[is.na(names$family),]

# no genus (and no family)
no_genus <- names[is.na(names$family),]

##########
# Family
saved=F
no_family_list <- list()
if(!saved){
for(i in 1: nrow(no_genus)){
    # powo
    r<- search_powo(no_genus$powo[i])
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      no_family_list[[i]] <- tmp
    } else {
      no_family_list[[i]] <- NA
    }
    write_rds(no_family_list, "../Extra/Names/no_family_list.R")
}
  } else {
  no_family_list_test <- read_rds("../Extra/Names/no_family_list.R")
  }

# only get all with data
no_family_name <- no_genus[!lengths(no_family_list)<2,]
no_family_list_all <- no_family_list
no_family_list <- no_family_list[!lengths(no_family_list)<2]


# get the correct and accepted names
family_safe <- data.frame(name = character(), family= character(), accepted= character())
for(i in 1: length(no_family_list)){
  dat <- no_family_list[[i]]
  acc_name <- NA
  acc <- NA
  name <- NA
  if(nrow(dat) == 1){
    multiple_safe <- rbind(family_safe, data.frame(name= no_family_name$powo[i], family = dat$family, accepted= dat$accepted))
  } else {
    acc_name <- dat$family[dat$accepted==T]
    freq <- table(acc_name)
    maxFreq <- which.max(freq)[1]   
    acc_name <-names(maxFreq)
    acc <- dat$accepted[dat$accepted==T]
    acc <- acc[!duplicated(acc)]
    name <- no_family_name$powo[i]
    family_safe <- rbind(family_safe, data.frame(name= name, family = acc_name, accepted= acc))
  }
}

names$family[is.na(names$family)] <- family_safe$family[match(names$powo[is.na(names$family)], family_safe$name)]

# check manyally
check <- names[is.na(names$family),] 
family_add <- c("Asteraceae","Asteraceae","Brassicaceae","Brassicaceae","Asteraceae","Asteraceae","Asteraceae","Ranunculaceae", NA, NA)
names$family[is.na(names$family)] <- family_add 
# wrong species names
wrong <- c("Brassica nigra", "Helleborus orientalis agg.")
names$powo[names$waardplant %in% wrong] <- c("Mutarda nigra","Helleborus orientalis")

#remove the errors
names <- names[!is.na(names$family),]
names <- names[!names$original=="Plantae indet.",]

##########
# no genus
no_genus <- names[is.na(names$genus),]
no_genus <- no_genus[!grepl("indet.",no_genus$original),]

##########
# Family
saved=F
no_family_list <- list()
if(!saved){
for(i in 1: nrow(no_genus)){
    # powo
    r<- search_powo(no_genus$powo[i])
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      no_family_list[[i]] <- tmp
    } else {
      no_family_list[[i]] <- NA
    }
    write_rds(no_family_list, "../Extra/Names/no_genus_list.R")
}
  } else {
  no_family_list_test <- read_rds("../Extra/Names/no_genus_list.R")
  }


# Genus
no_genus_list <- no_family_list

# only get all with data
no_genus_name <- no_genus[!lengths(no_genus_list)<2,]
no_genus_list_all <- no_genus_list
no_genus_list <- no_genus_list[!lengths(no_genus_list)<2]
i=2
# get the correct and accepted names
genus_safe <- data.frame(name = character(), genus= character(), accepted= character())
for(i in 1: length(no_genus_list)){
  dat <- no_genus_list[[i]]
  acc_name <- NA
  acc <- NA
  name <- NA
  if(nrow(dat) == 1){
    acc <- ifelse(dat$rank=="Genus", dat$name, NA)
    genus_safe <- rbind(genus_safe, data.frame(name= no_genus_name$powo[i], genus = acc, accepted= dat$accepted))
  } else {
    acc_name <- dat$name[dat$rank=="Genus" & dat$accepted=="TRUE"]
    freq <- table(acc_name)
    maxFreq <- which.max(freq)[1]   
    acc_name <-names(maxFreq)
    acc <- dat$accepted[dat$rank=="Genus" & dat$accepted=="TRUE"]
    acc <- acc[!duplicated(acc)]
    acc_name <- ifelse(is.null(acc_name), dat$name[dat$rank=="Genus"], acc_name)
    acc <- ifelse(length(acc)<1, dat$accepted[dat$rank=="Genus"], acc)
    name <- no_genus_name$powo[i]
    genus_safe <- rbind(genus_safe, data.frame(name= name, genus = acc_name, accepted= acc))
  }
}

# assign
names$genus[is.na(names$genus)] <- genus_safe$genus[match(names$powo[is.na(names$genus)], genus_safe$name)]


# check manyally
no_genus <- names[is.na(names$genus),] 
no_genus <- no_genus[!grepl("indet.",no_genus$original),]

no_genus$genus <- c(NA,"× Brassarda","Mutarda",NA,NA, "Anacis","Helleborus")
names$genus[is.na(names$genus) & names$original %in% no_genus$original] <- 
  no_genus$genus[match(names$original[is.na(names$genus)& names$original %in% no_genus$original] , no_genus$original)]
```


Check remaining names
```{r}
# some changes
names$genus[names$waardplant == names$family] <- NA
names$genus[names$waardplant =="Bambusoideae"] <- NA
names[names$waardplant=="Lotus glaber", 4:5] <- data.frame("Lotus tenuis","Lotus")
names[names$waardplant=="Aronia", 5] <- data.frame("Aronia")
names[names$waardplant=="Stellaria", 5] <- data.frame("Stellaria")
names[names$waardplant=="Pastinaca", 5:6] <- data.frame("Pastinaca","Apiaceae")
names[names$waardplant=="Anchusa arvensis", 4:5] <- data.frame("Anchusa arvensis","Anchusa")
names[names$waardplant=="Trifolium", 5] <- data.frame("Trifolium")
names[names$waardplant=="Fallopia japonica + Fallopia sachalinensis + Fallopia x bohemica", 4:5] <- data.frame("Reynoutria × bohemica", "Reynoutria")
names[names$waardplant=="Eriobotrya japonica", 4:5] <- data.frame("Eriobotrya japonica", "Eriobotrya")
names[names$waardplant=="Perovskia ", 4] <- data.frame("Salvia")
names[names$waardplant=="Lychnis", 4] <- data.frame("Silene")
names[names$waardplant=="Silene armeria", 5] <- data.frame("Silene")
names[names$powo=="Salix alba", 5] <- data.frame("Salix")
names[names$powo=="Cnicus", 5] <- data.frame("Centaurea")
names[names$powo=="Convolvulus soldanella", 4:5] <- data.frame("Calystegia soldanella","Calystegia")
names[names$powo=="Inula conyzae", 4:5] <- data.frame("Pentanema squarrosum","Pentanema")
names[names$powo=="Leontodon saxatilis", 5] <- data.frame("Leontodon")
names[names$powo=="Peucedanum palustre", 5] <- data.frame("Peucedanum")
names[names$powo=="Potentilla rupestris", 4:5] <- data.frame("Drymocallis rupestris","Drymocallis")
names[names$powo=="Sedum sediforme", 4:5] <- data.frame("Petrosedum sediforme","Petrosedum")
names[names$powo=="Sedum rupestre", 4:5] <- data.frame("Petrosedum rupestre","Petrosedum")
names[names$powo=="Sorbus intermedia", 4:5] <- data.frame("Scandosorbus intermedia","Scandosorbus")
names[names$powo=="Brasica and Sinapis", 6] <- data.frame("Brassicaceae")
names[names$original =="Sinapis arvensis", 4:5] <- data.frame("Mutarda arvensis", "Mutarda")
```




Reduce dataset
```{r}
# which do I remove
remove_waardplanten <- polplant[!polplant$waardplant %in% names$original,]
# which do I keep
polplant <- polplant[polplant$waardplant %in% names$original,]
# give accepted name
polplant$powo <- names$powo[match(polplant$waardplant, names$original)]
polplant$genus_plants <- names$genus[match(polplant$waardplant, names$original)] 
polplant$family_plants <- names$family[match(polplant$waardplant, names$original)]
polplant$plant_level <- ifelse(is.na(polplant$genus_plants), "Family",NA)
polplant$plant_level[is.na(polplant$plant_level)] <- ifelse(polplant$genus_plants[is.na(polplant$plant_level)]==
                                                              polplant$powo[is.na(polplant$plant_level)], "Genus","Species")
```

Remove some data
```{r}
rm(list = setdiff(ls(),c("names", "remove_waardplanten", "polplant")))
write_csv(names, "../Extra/Names/plant_names.csv")
```


## 2.2 Pollinators
```{r}
# this has to be done against the gbif data
pol <- polplant |> group_by(naam_lat, naam_nl) |> summarise(n=n())

# gbif
gbif <- name_backbone_checklist(pol$naam_lat)

# assign gbif names
pol$gbif <- gbif$species
pol$genus <- gbif$genus
pol$family <- gbif$family

# some are different, check
different_name <- pol[pol$naam_lat != pol$gbif,]
# check dupplicates -> use gbif name
dup <- pol[duplicated(pol$gbif) | duplicated(pol$gbif, fromLast = T),]
# check remainings
# different_name <- different_name[!different_name$naam_lat %in% dup$naam_lat,]
# assign the original name instead of gbif name
pol$gbif[pol$naam_lat %in% different_name$naam_lat] <- pol$naam_lat[pol$naam_lat %in% different_name$naam_lat]
# adapt the aardhommelcomplex
pol$gbif[pol$naam_nl=="aardhommelcomplex"] <- "Bombus terrestris complex"
```



Assign to database
```{r}
polplant$gbif <- pol$gbif[match(polplant$naam_lat, pol$naam_lat)]
polplant$genus_pol <- pol$genus[match(polplant$naam_lat, pol$naam_lat)]
polplant$family_pol <- pol$family[match(polplant$naam_lat, pol$naam_lat)]
rm(list = setdiff(ls(),c("names", "remove_waardplanten", "polplant", "pol")))
```


## 2.3 Pictures
```{r}
# load data
pictures_polplant <- read.csv2("../Data/2024-145_BE_metWP_sws+metToestemming_linkFotoID.csv", sep= ",")

# assign names
pictures_polplant$species <- polplant$gbif[match(pictures_polplant$wnm_id, polplant$id)]
sum(!duplicated(pictures_polplant$wnm_id))

# check all removed species
removed <- pictures_polplant[(is.na(pictures_polplant$species)),]
removed$species <- remove_waardplanten$waardplant[match(removed$wnm_id, remove_waardplanten$id)]

# how many duplicates are there in the database
# 42% of all observations are unique, so many duplicates
length(unique(pictures_polplant$wnm_id))/ length(pictures_polplant$foto_id)

# assign link picture to polplant
polplant$picture_id <- pictures_polplant$foto_id[match(polplant$id, pictures_polplant$wnm_id)]
sum(!is.na(polplant$picture_id))
# check how many of the species have pictures
# about 20% of the observations have images
sum(!is.na(polplant$picture_id))/ length(polplant$id)

#write_csv(pictures_polplant, "../Data/polplant_pictures.csv")
```



## 2.4 Header
General
```{r}
# check names
colnames(polplant)
#str(polplant)

# remove some unneccessary columns
polplant <- select(polplant, -utm1, -utm5, -bijzonderheden, -aanleverdatum)

#change column names
colnames(polplant) <- c("ID","Name","Name_nl","Date","Type","SpeciesID","Number","Behaviour","Method","Stage","Gender","PlantID","Plant","Status",
                        "Number_pictures","x","y","Long", "Lat","powo","genus_plants","family_plants","plant_level","gbif","genus_pol","family_pol","PictureID")

# change order
polplant <- polplant |> relocate("gbif", .before="Name")
polplant <- polplant |> relocate("powo", .before="Plant")
polplant <- polplant |> relocate("PictureID", .after="ID")

# check if any (except for the pictures and genera of plants) is NA
any(is.na(polplant[,c(1,3:22, 24:26)]))
```


Time
```{r}
# change dates
polplant$Date <- as.Date(polplant$Date, format= c("%d/%m/%Y"))

# plot
ggplot(polplant)+
  geom_histogram(aes(x=Date), fill="lightgrey")+ scale_y_log10()+
  theme_pubr()+
  # Observations with a picture (blue overlay)
  geom_histogram(data = subset(polplant, !is.na(PictureID)),
                 aes(x = Date), fill = "steelblue", alpha = 1) +
  ylab("Number of observations")+
  xlab("Year")
```

Coordinates
```{r}
# Convert coordinates to sf points
coord <- polplant[, c("ID", "Long", "Lat")] %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

# Get Belgium polygon as sf
Belgium_df <- map_data("world") %>% filter(region == "Belgium")
Belgium_sf <- st_as_sf(Belgium_df, coords = c("long", "lat"), crs = 4326) %>%
  group_by(group) %>% summarise(geometry = st_combine(geometry)) %>%  st_cast("POLYGON") %>%st_union() %>%st_sf()

# the whole world
world_df <- map_data("world")
world_sf <- st_as_sf(world_df, coords = c("long", "lat"), crs = 4326) %>%
  group_by(group) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")
bbox <- st_as_sfc(st_bbox(Belgium_sf))
world_clipped <- st_crop(world_sf, xmin = 2.5, xmax = 6.5, ymin = 49.4, ymax = 51.6)

# Create 0.05° grid and intersect with Belgium
grid <- st_make_grid(Belgium_sf, cellsize = 0.05, square = TRUE) %>%
  st_sf() %>%
  st_intersection(Belgium_sf)

# Count points per grid cell
coord$obs <- 1
grid_counts <- st_join(grid, coord, join = st_intersects) %>%
  group_by(geometry) %>%
  summarise(n = sum(obs, na.rm = TRUE))  # NA if no points

grid_counts$n[grid_counts$n=="0"] <- NA
grid_counts$obs_cat <- ifelse(is.na(grid_counts$n), "No data", "Data")
grid_counts$log_n <- log10(grid_counts$n)

ggplot() +
  # Background: clipped world map
  #geom_sf(data = world_clipped, fill = "antiquewhite", colour = NA) +
  # Country outlines (actual borders only)
  #geom_path(data = world_df, aes(x = long, y = lat, group = group),colour = "grey80", size = 0.2)+
  # Belgium fill for contrast
  geom_sf(data = Belgium_sf, fill = "antiquewhite", colour = "black", size = 0.3) +
  # Grid cells with data (log10 scale)
  geom_sf(data = filter(grid_counts, obs_cat == "Data"),
          aes(fill = log_n), colour = NA) +
  # Grid cells with no data (antiquewhite with grey frame)
  geom_sf(data = filter(grid_counts, obs_cat == "No data"),
          fill = "antiquewhite", colour = "antiquewhite", size = 0.2) +
  # Continuous fill scale
  scale_fill_gradientn(
    colours = viridis::viridis(100, option = "plasma"),
    breaks = c(0, 1, 2, 3, 4),
    labels = c("1", "10", "100", "1,000", "10,000"),
    name = "Observations",
    na.value = 0
  ) +
  # Add dummy legend for NA values
  new_scale_fill() + 
  geom_sf(data = filter(grid_counts, obs_cat == "No data"),
          aes(fill = obs_cat), show.legend = TRUE, colour = NA) +
  scale_fill_manual(    
    guide = guide_legend(override.aes = list(colour = "antiquewhite")),
    name = NULL,
    values = c("No data" = "antiquewhite")
  ) +
  coord_sf(xlim = c(2.5, 6.5), ylim = c(49.4, 51.6), expand = FALSE) +
  theme_pubclean() +
  theme(legend.position = "right") +
  labs(x = "Longitude", y = "Latitude")

#ggsave("../Images/map_observations_belgium.png", width= 7, height= 7)
```

Categories
```{r}
# check observation categories
polplant |> group_by(Type) |> summarise(n=n())
# explantation:
# blauw (accepted by admin), grijs (accepted by image recognition), groen (accepted with evidence)
```


```{r}
rm(list = setdiff(ls(),c("names", "remove_waardplanten", "polplant", "pol")))
```


## 2.5 Origins
```{r}
origin <- polplant |> group_by(powo,Status) |> summarise(n=length(unique(powo)))

# check aliens
Belgium <- read_csv( "../Extra/GLONAF/Belgium.csv", show_col_types = FALSE)
origin$glonaf <- Belgium$status[match(origin$powo, Belgium$taxa_accepted)]

# easier, just alien and native
polplant$Status[polplant$Status %in% c("Incidental import", "Naturalized","Does not occur")] <- "Alien"
polplant$Status[polplant$Status %in% c("Expected","Native","Extinct")] <- "Native"
unique(polplant$Status)
```

```{r}
status <- read_csv("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/GLONAF/Belgium.csv", show_col_types = FALSE)
unique(status$status)
names$source <- NA
# assign Belgium
names$status <- status$status[match(names$powo, status$taxon_corrected)]
names$source[!is.na(names$status)] <- "GloNAF"
sum(!is.na(names$status))
```

```{r}
status <- readxl::read_excel("C:/Users/u0166342/Documents/Doctoraat/Impact/Extra data/Species names/Veronika_name_correction.xlsx")
extra <- status[status$Status.checked.VK =="extra",]
extra <- extra[!duplicated(extra$`Aggregated name`),]
status <- status[status$Region=="Belgium",]

# assign Belgium
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$powo[is.na(names$status)], status$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$gbif[is.na(names$status)] , status$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$original[is.na(names$status)] , status$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$powo[is.na(names$status)] , status$`Original EVA name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$gbif[is.na(names$status)] , status$`Original EVA name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$original[is.na(names$status)] , status$`Original EVA name`)]
sum(!is.na(names$status))

# assign extra-European
names$status[is.na(names$status)] <- extra$Status.checked.VK[match(names$powo[is.na(names$status)] , extra$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$Status.checked.VK[match(names$original[is.na(names$status)] , extra$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$Status.checked.VK[match(names$powo[is.na(names$status)] , extra$`Original EVA name`)]
sum(!is.na(names$status))

names$source[!(is.na(names$status)) & is.na(names$source)] <- "Kalusova"
```


```{r}
status <- readxl::read_excel("C:/Users/u0166342/Documents/Doctoraat/Impact/Extra data/Species names/Irena_taxa_update.xlsx")
extra <- status[status$statusNew =="extra",]
extra <- extra[!duplicated(extra$species),]

status <- status[status$Region=="Belgium",]


# assign Belgium
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)], status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))

# assign extra-European
names$status[is.na(names$status)] <- extra$statusNew[match(names$powo[is.na(names$status)] , extra$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$statusNew[match(names$original[is.na(names$status)] , extra$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$statusNew[match(names$powo[is.na(names$status)] , extra$`Matched concept`)]
sum(!is.na(names$status))

names$source[!(is.na(names$status)) & is.na(names$source)] <- "Kalusova"
```





```{r}
species_mab <- read.delim("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/Names/Aliens/taxon.txt")
status_mab <- read.delim("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/Names/Aliens/distribution.txt")

species_mab$species <- vegdata::taxname.removeAuthors(species_mab$scientificName)

status<- species_mab
status$statusNew <- "Alien"

# assign Belgium
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)], status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))

names$source[!(is.na(names$status)) & is.na(names$source)] <- "Manual alien plants Belgium"

```



```{r}
names$status[names$status %in% c("Naturalized","Invasive","intra","extra")] <- "Alien"
names$status[names$status %in% c("native")] <- "Native"


check <- names[!is.na(names$status),]

# remaining
remaining <-names[is.na(names$status),]
remaining$waarnemingen <- polplant$Status[match(remaining$original, polplant$Plant)]
remaining_aliens <- remaining[remaining$waarnemingen =="Alien",]
# I checked some, all were indeed alien

# Assign
names$source[is.na(names$source)] <- "Waarnemingen"
names$status[is.na(names$status)] <- remaining$waarnemingen
```

```{r}
status <- read_csv("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/GLONAF/Belgium.csv", show_col_types = FALSE)
status <- status[status$status=="Invasive",]

names$invasive <- status$status[match(names$powo, status$taxon_corrected)]
sum(!is.na(names$invasive))
names$invasive[is.na(names$invasive)] <- status$status[match(names$gbif[is.na(names$invasive)], status$taxon_corrected)]
sum(!is.na(names$invasive))
names$invasive[is.na(names$invasive)] <- status$status[match(names$waardplant[is.na(names$invasive)], status$taxon_corrected)]
sum(!is.na(names$invasive))
names$invasive[is.na(names$invasive)] <- status$status[match(names$original[is.na(names$invasive)], status$taxon_corrected)]
sum(!is.na(names$invasive))

# based on the invasive alien species in Belgium document
species_vector_invasives <- c(
  "Acer negundo",
  "Acer rufinerve",
  "Ailanthus altissima",
  "Akebia quinata",
  "Ambrosia artemisiifolia",
  "Amelanchier lamarckii",
  "Aster lanceolatus",
  "Aster novi-belgii",
  "Aster x salignus",
  "Azolla filiculoides",
  "Baccharis halimifolia",
  "Bidens frondosa",
  "Buddleja davidii",
  "Cabomba caroliniana",
  "Carpobrotus spp.",
  "Cornus sericea",
  "Cotoneaster horizontalis",
  "Crassula helmsii",
  "Cyperus eragrostis",
  "Duchesnea indica",
  "Echinocystis lobata",
  "Egeria densa",
  "Elaeagnus angustifolia",
  "Elodea canadensis",
  "Elodea nuttallii",
  "Epilobium ciliatum",
  "Fallopia japonica",
  "Fallopia sachalinensis",
  "Fallopia x bohemica",
  "Fraxinus pennsylvanica",
  "Gaillardia x grandiflora",
  "Helianthus tuberosus",
  "Heracleum mantegazzianum",
  "Hyacinthoides hispanica",
  "Hydrocotyle ranunculoides",
  "Impatiens glandulifera",
  "Impatiens parviflora",
  "Koenigia polystachya",
  "Lagarosiphon major",
  "Lemna minuta",
  "Lonicera japonica",
  "Ludwigia grandiflora",
  "Ludwigia peploides",
  "Lupinus polyphyllus",
  "Lysichiton americanus",
  "Mahonia aquifolium",
  "Mimulus guttatus",
  "Myriophyllum aquaticum",
  "Myriophyllum heterophyllum",
  "Oenothera spp.",
  "Parthenocissus spp.",
  "Phytolacca americana",
  "Prunus laurocerasus",
  "Prunus serotina",
  "Quercus rubra",
  "Rhododendron ponticum",
  "Rhus typhina",
  "Robinia pseudoacacia",
  "Rosa rugosa",
  "Rudbeckia laciniata",
  "Senecio inaequidens",
  "Solidago canadensis",
  "Solidago gigantea",
  "Spiraea alba",
  "Spiraea douglasii",
  "Spiraea tomentosa",
  "Spiraea x billardii"
)

names$invasive[is.na(names$invasive) & names$powo %in% species_vector_invasives] <- "Invasive"

sum(!is.na(names$invasive))

#write_csv(names, "../Extra/Names/Estelle_origins.csv")
```



# 3 Basic numbers
## 3.1 Plot
Calculate numbers
```{r}
# number of pollinator species
num_pol <- length(unique(polplant$gbif))

# number of plant species
num_plant <- length(unique(polplant$powo))

# number of plant species per pollinator
num_rel_species <- polplant |> group_by(gbif) |> summarise(n = length(unique(powo)), rel = n/num_plant)
num_rel_genus <- polplant |> group_by(genus_pol) |> summarise(n = length(unique(powo)), rel = n/num_plant)
num_rel_family <- polplant |> group_by(family_pol) |> summarise(n = length(unique(powo)), rel = n/num_plant)

# number of pollinators per plant
plant_num_rel_species <- polplant |> group_by(powo) |> summarise(n = length(unique(gbif)), rel = n/num_pol)
plant_num_rel_genus <- polplant |> group_by(genus_plants) |> summarise(n = length(unique(gbif)), rel = n/num_pol)
plant_num_rel_family <- polplant |> group_by(family_plants) |> summarise(n = length(unique(gbif)), rel = n/num_pol)
```

Some basic plots
```{r}
# number of associated species
ggplot()+
  geom_histogram(data=num_rel_species, aes(x= n), color= NA, fill="darkblue")+
  theme_classic()
```

```{r}
# Create a data frame with count of interactions
links <- polplant %>%
  count(family_pol, family_plants, name = "value")

links <- links[links$value >29,]

# Force order: pollinator families first, then plant families
poll_families <- unique(links$family_pol)
plant_families <- unique(links$family_plants)

nodes <- data.frame(name = c(poll_families, plant_families))

# Recode links to use 0-based indexing
links$source <- match(links$family_pol, nodes$name) - 1
links$target <- match(links$family_plants, nodes$name) - 1

# Add grouping for colouring (optional)
nodes$group <- ifelse(nodes$name %in% poll_families, "Pollinator", "Plant")

# Create Sankey diagram
sankeyNetwork(Links = links, Nodes = nodes,
              Source = "source", Target = "target",
              Value = "value", NodeID = "name",
              NodeGroup = NULL, 
              fontSize = 12, nodeWidth = 30,
              sinksRight = TRUE)
```

Interaction matrix
```{r}
# create matrix
interaction_matrix <- polplant |>dplyr::count(family_plants, family_pol) |> tidyr::pivot_wider(names_from = family_pol,values_from = n,values_fill = list(n = 0))
interaction_matrix <- as.data.frame(interaction_matrix)
rownames(interaction_matrix) <- interaction_matrix$family_plants
interaction_matrix <- interaction_matrix[, -1]
# plot
plotweb(interaction_matrix, text.rot =90)
```


## 3.2 Origins
```{r}
num_rel_origin <- polplant |> group_by(powo,Status) |> summarise(unique=length(unique(gbif)), n=n())

ggplot(data=num_rel_origin, aes(x= Status, y=unique, fill= Status))+
  geom_violin(alpha= 0.5, scale= "width")+
  geom_boxplot(width= 0.25, alpha=0.8, fill="white")+
  theme_classic()+
  ylab("Number of associated pollinators")+
  xlab("Species status")+
  theme(legend.position = "none")+
  scale_colour_manual(values=c("#1E88E5",  "#004D40"))+
  scale_fill_manual(values=c("#1E88E5",  "#004D40"))+
  stat_summary(fun= "mean",
               geom = "point", aes(group= Status), size=3)+
  ggpubr::stat_compare_means(comparisons = list(c("Alien","Native")), label="p.signif", method= "t.test")+
  scale_y_log10()
```



# POL
# 1 START
Remove old data
```{r}
pol_names <- pol
rm(list = setdiff(ls(),c("names", "remove_waardplanten", "polplant", "pol_names")))
```

Download pollinator only data
```{r}
poldata <- read.csv2("../Data/2024-145_BE_zonderWP_sws+metToestemming.csv")
```


# 2 Data cleaning
## 2.1 Pollinators
```{r}
# this has to be done against the gbif data
pol <- poldata |> group_by(naam_lat, naam_nl) |> summarise(n=n())

# gbif
gbif <- name_backbone_checklist(pol$naam_lat)

# assign gbif names
pol$gbif <- gbif$species
pol$genus <- gbif$genus
pol$family <- gbif$family

# some are different, check
different_name <- pol[pol$naam_lat != pol$gbif,]
# check dupplicates -> use original name (dutch names differ)
dup <- pol[duplicated(pol$gbif) | duplicated(pol$gbif, fromLast = T),]
# check remainings
# different_name <- different_name[!different_name$naam_lat %in% dup$naam_lat,]
# assign the original name instead of gbif name
pol$gbif[pol$naam_lat %in% different_name$naam_lat] <- pol$naam_lat[pol$naam_lat %in% different_name$naam_lat]
# adapt the aardhommelcomplex
pol$gbif[pol$naam_nl=="aardhommelcomplex"] <- "Bombus terrestris complex"

# two species were not found in gbif
not_found <- pol[is.na(pol$gbif),]
not_found$gbif <- not_found$naam_lat
not_found$genus <- "Stelis"
not_found$family <- "Megachilidae"
pol[is.na(pol$gbif),] <- not_found 
```


Assign to database
```{r}
poldata$gbif <- pol$gbif[match(poldata$naam_lat, pol$naam_lat)]
poldata$genus_pol <- pol$genus[match(poldata$naam_lat, pol$naam_lat)]
poldata$family_pol <- pol$family[match(poldata$naam_lat, pol$naam_lat)]
rm(list = setdiff(ls(),c("names", "remove_waardplanten", "poldata", "pol", "polplant", "pol_names")))
```



## 2.2 Pictures
```{r}
# load data
pictures_poldata <- read.csv2("../Data/2024-145_BE_zonderWP_sws+metToestemming_linkFotoID.csv", sep= ",")

# assign names
pictures_poldata$species <- poldata$gbif[match(pictures_poldata$wnm_id, poldata$id)]

# how many duplicates are there in the database
# 68% of all observations are unique, so many duplicates
length(unique(pictures_poldata$wnm_id))/ length(pictures_poldata$foto_id)

# assign link picture to poldata
poldata$picture_id <- pictures_poldata$foto_id[match(poldata$id, pictures_poldata$wnm_id)]
# check how many of the species have pictures
# about 75% of the observations have images
sum(!is.na(poldata$picture_id))/ length(poldata$id)

write_csv(pictures_poldata, "../Data/poldata_pictures.csv")
```


## 2.3 Header
General
```{r}
# check names
colnames(poldata)
#str(poldata)

# remove some unneccessary columns
poldata <- select(poldata, -utm1, -utm5, -bijzonderheden, -aanleverdatum)

#change column names
colnames(poldata) <- c("ID","Name","Name_nl","Date","Type","SpeciesID","Number","Behaviour","Method","Stage","Gender",
                        "Number_pictures","x","y","Long", "Lat","gbif","genus_pol","family_pol", "PictureID")

# change order
poldata <- poldata |> relocate("gbif", .before="Name")
poldata <- poldata |> relocate("PictureID", .after="ID")

# check if any (except for the pictures) is NA
any(is.na(poldata[,c(1,3:20)]))
```


Time
```{r}
# change dates
poldata$Date <- as.Date(poldata$Date, format= c("%d/%m/%Y"))

# plot
ggplot(poldata)+
  geom_histogram(aes(x=Date), fill="lightgrey")+ scale_y_log10()+
  theme_pubr()+
  # Observations with a picture (blue overlay)
  geom_histogram(data = subset(poldata, !is.na(PictureID)),
                 aes(x = Date), fill = "steelblue", alpha = 1) +
  ylab("Number of observations")+
  xlab("Year")
```



Coordinates
```{r}
# Convert coordinates to sf points
coord <- poldata[, c("ID", "Long", "Lat")] %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

# Get Belgium polygon as sf
Belgium_df <- map_data("world") %>% filter(region == "Belgium")
Belgium_sf <- st_as_sf(Belgium_df, coords = c("long", "lat"), crs = 4326) %>%
  group_by(group) %>% summarise(geometry = st_combine(geometry)) %>%  st_cast("POLYGON") %>%st_union() %>%st_sf()

# the whole world
world_df <- map_data("world")
world_sf <- st_as_sf(world_df, coords = c("long", "lat"), crs = 4326) %>%
  group_by(group) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")
bbox <- st_as_sfc(st_bbox(Belgium_sf))
world_clipped <- st_crop(world_sf, xmin = 2.5, xmax = 6.5, ymin = 49.4, ymax = 51.6)

# Create 0.05° grid and intersect with Belgium
grid <- st_make_grid(Belgium_sf, cellsize = 0.05, square = TRUE) %>%
  st_sf() %>%
  st_intersection(Belgium_sf)

# Count points per grid cell
coord$obs <- 1
grid_counts <- st_join(grid, coord, join = st_intersects) %>%
  group_by(geometry) %>%
  summarise(n = sum(obs, na.rm = TRUE))  # NA if no points

grid_counts$n[grid_counts$n=="0"] <- NA
grid_counts$obs_cat <- ifelse(is.na(grid_counts$n), "No data", "Data")
grid_counts$log_n <- log10(grid_counts$n)

ggplot() +
  # Background: clipped world map
  #geom_sf(data = world_clipped, fill = "antiquewhite", colour = NA) +
  # Country outlines (actual borders only)
  #geom_path(data = world_df, aes(x = long, y = lat, group = group),colour = "grey80", size = 0.2)+
  # Belgium fill for contrast
  geom_sf(data = Belgium_sf, fill = "antiquewhite", colour = "black", size = 0.3) +
  # Grid cells with data (log10 scale)
  geom_sf(data = filter(grid_counts, obs_cat == "Data"),
          aes(fill = log_n), colour = NA) +
  # Grid cells with no data (antiquewhite with grey frame)
  geom_sf(data = filter(grid_counts, obs_cat == "No data"),
          fill = "antiquewhite", colour = "antiquewhite", size = 0.2) +
  # Continuous fill scale
  scale_fill_gradientn(
    colours = viridis::viridis(100, option = "plasma"),
    breaks = c(0, 1, 2, 3, 4),
    labels = c("1", "10", "100", "1,000", "10,000"),
    name = "Observations",
    na.value = 0
  ) +
  # Add dummy legend for NA values
  new_scale_fill() + 
  geom_sf(data = filter(grid_counts, obs_cat == "No data"),
          aes(fill = obs_cat), show.legend = TRUE, colour = NA) +
  scale_fill_manual(    
    guide = guide_legend(override.aes = list(colour = "antiquewhite")),
    name = NULL,
    values = c("No data" = "antiquewhite")
  ) +
  coord_sf(xlim = c(2.5, 6.5), ylim = c(49.4, 51.6), expand = FALSE) +
  theme_pubclean() +
  theme(legend.position = "right") +
  labs(x = "Longitude", y = "Latitude")

#ggsave("../Images/map_observations_pol_only_belgium.png", width= 7, height= 7)
```


Categories
```{r}
# check observation categories
poldata |> group_by(Type) |> summarise(n=n())
# explantation:
# blauw (accepted by admin), grijs (accepted by image recognition), groen (accepted with evidence)
```


```{r}
rm(list = setdiff(ls(),c("names", "remove_waardplanten", "polplant", "pol", "poldata", "pol_names")))
```

# SAVE
```{r}
write_csv(polplant,"../Data/polplant.csv")
#write_csv(poldata,"../Data/poldata.csv")
```


# COMBINE
# 1 Datasets
```{r}
colnames(poldata)
poldata$dataset <- "new"
colnames(polplant[, c(1:13,18:22, 26:27)])
polplant$dataset <- "val"
all <- rbind(poldata, polplant[, c(1:13,18:22, 26:28)])
```


```{r}
# Convert coordinates to sf points
coord <- all[, c("ID", "Long", "Lat")] %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

# Get Belgium polygon as sf
Belgium_df <- map_data("world") %>% filter(region == "Belgium")
Belgium_sf <- st_as_sf(Belgium_df, coords = c("long", "lat"), crs = 4326) %>%
  group_by(group) %>% summarise(geometry = st_combine(geometry)) %>%  st_cast("POLYGON") %>%st_union() %>%st_sf()

# the whole world
world_df <- map_data("world")
world_sf <- st_as_sf(world_df, coords = c("long", "lat"), crs = 4326) %>%
  group_by(group) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")
bbox <- st_as_sfc(st_bbox(Belgium_sf))
world_clipped <- st_crop(world_sf, xmin = 2.5, xmax = 6.5, ymin = 49.4, ymax = 51.6)

# Create 0.05° grid and intersect with Belgium
grid <- st_make_grid(Belgium_sf, cellsize = 0.05, square = TRUE) %>%
  st_sf() %>%
  st_intersection(Belgium_sf)

# Count points per grid cell
coord$obs <- 1
grid_counts <- st_join(grid, coord, join = st_intersects) %>%
  group_by(geometry) %>%
  summarise(n = sum(obs, na.rm = TRUE))  # NA if no points

grid_counts$n[grid_counts$n=="0"] <- NA
grid_counts$obs_cat <- ifelse(is.na(grid_counts$n), "No data", "Data")
grid_counts$log_n <- log10(grid_counts$n)

ggplot() +
  # Background: clipped world map
  #geom_sf(data = world_clipped, fill = "antiquewhite", colour = NA) +
  # Country outlines (actual borders only)
  #geom_path(data = world_df, aes(x = long, y = lat, group = group),colour = "grey80", size = 0.2)+
  # Belgium fill for contrast
  geom_sf(data = Belgium_sf, fill = "antiquewhite", colour = "black", size = 0.3) +
  # Grid cells with data (log10 scale)
  geom_sf(data = filter(grid_counts, obs_cat == "Data"),
          aes(fill = log_n), colour = NA) +
  # Grid cells with no data (antiquewhite with grey frame)
  geom_sf(data = filter(grid_counts, obs_cat == "No data"),
          fill = "antiquewhite", colour = "antiquewhite", size = 0.2) +
  # Continuous fill scale
  scale_fill_gradientn(
    colours = viridis::viridis(100, option = "plasma"),
    breaks = c(0, 1, 2, 3, 4),
    labels = c("1", "10", "100", "1,000", "10,000"),
    name = "Observations",
    na.value = 0
  ) +
  # Add dummy legend for NA values
  new_scale_fill() + 
  geom_sf(data = filter(grid_counts, obs_cat == "No data"),
          aes(fill = obs_cat), show.legend = TRUE, colour = NA) +
  scale_fill_manual(    
    guide = guide_legend(override.aes = list(colour = "antiquewhite")),
    name = NULL,
    values = c("No data" = "antiquewhite")
  ) +
  coord_sf(xlim = c(2.5, 6.5), ylim = c(49.4, 51.6), expand = FALSE) +
  theme_pubclean() +
  theme(legend.position = "right") +
  labs(x = "Longitude", y = "Latitude")

#ggsave("../Images/map_observations_all_belgium.png", width= 7, height= 7)
```


```{r}
# change to date
all$Date <- as.Date(all$Date, format= c("%d/%m/%Y"))

# plot
ggplot(all)+
  geom_histogram(aes(x=Date), fill="lightgrey")+ scale_y_log10()+
  theme_pubr()+
  # Observations with a picture (blue overlay)
  geom_histogram(data = subset(all, !is.na(PictureID)),
                 aes(x = Date), fill = "steelblue", alpha = 1) +
  geom_histogram(data = all[all$dataset=="val",],
             aes(x = Date), fill = "darkblue", alpha = 1)+
  ylab("Number of observations")+
  xlab("Year")
  

ggsave("../Images/all_observations_time.png", width= 7, height= 5)
```


# ESTELLE
## Names
```{r}
plants <- read_csv("C:/Users/u0166342/OneDrive - KU Leuven/Doctoraat/Teaching assistant/Thesis/2024-2025/Estelle/Data/list_plant_status.csv")

names <- unique(plants[,1:4])
names$waardplant <- gsub("\\b(spec\\.)\\b", "", names$original_name)
names$waardplant <- trimws(names$waardplant)
indet <- names$waardplant[grepl(" indet.", names$waardplant)]
names$waardplant <- gsub(" indet.","", names$waardplant)

gbif <- name_backbone_checklist(names$waardplant)

# check which are not plant in gbif
non_plants <- gbif[!gbif$kingdom =="Plantae",]
non_plants <- non_plants[!is.na(non_plants$verbatim_name),]

# check in powo 
names_powo <- match_knms(names$waardplant)
names_powo <- names_powo$results

# safe depending on types
safe_names <- c()
multiple <- c()
none <- c()
# results can be false true multiple_matches or ""
# all together also 1697
for(i in 1: length(names_powo)){
  if(names_powo[[i]][[2]]== "true"){
    safe_names <- rbind(safe_names, names_powo[[i]][[1]])
  } else if(names_powo[[i]][[2]]=="multiple_matches"){
    multiple <- rbind(multiple, names_powo[[i]][[1]])
  } else if(names_powo[[i]][[2]]=="false"){
    none <- rbind(none, names_powo[[i]][[1]])
  }
}

################
# SAFE
# give powo names also to name list
names$powo <- safe_names[match(names$waardplant, safe_names)]
names$genus <- gbif$genus[match(names$powo, gbif$canonicalName)]
names$family <- gbif$family[match(names$powo, gbif$canonicalName)]



################
# NONE
# search powo and store results in list if we find a direct match
none_powo <- list()
saved = F
if(!saved){
for(i in 1: length(none)){
    # powo
    r<- search_powo(none[i])
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      none_powo[[i]] <- tmp
    } else {
      none_powo[[i]] <- NA
    }
    #write_rds(none_powo, "../Extra/Names/powo_no_link.R")
}
} else {
  none_powo <- read_rds("../Extra/Names/powo_no_link.R")
}

# only get all with data
none_name <- none[!lengths(none_powo)<2]
none_powo_all <- none_powo
none_powo <- none_powo[!lengths(none_powo)<2]

# get the correct and accepted names
none_safe <- data.frame(old = character(), new= character(), accepted= character())
for(i in 1: length(none_powo)){
  dat <- none_powo[[i]]
  acc_name <- NA
  acc <- NA
  old <- NA
  if(nrow(dat) == 1){
    none_safe <- rbind(none_safe, data.frame(old= none_name[i], new = dat$name, accepted= dat$accepted))
  } else {
    acc_name <- dat$name[dat$accepted==T]
    acc <- dat$accepted[dat$accepted==T]
    old <- none_name[i]
    none_safe <- rbind(none_safe, data.frame(old= old, new = acc_name, accepted= acc))
  }
}

dup <- none_safe[duplicated(none_safe$old)| duplicated(none_safe$old, fromLast=T),]
remove <- dup[c(1,2),]
none_safe <- none_safe[!(none_safe$old %in% remove$old & none_safe$new %in% remove$new),]

none_safe[none_safe$old %in% c("Astereae","Eupatorieae","Heliantheae"), 2:3] <- NA
names$powo[is.na(names$powo)] <- none_safe$new[match(names$waardplant[is.na(names$powo)], none_safe$old)]


################
# MULTIPLE
# search powo and store results in list if we find a direct match
multiple_powo <- list()
if(!saved){
for(i in 1: length(multiple)){
    # powo
    r<- search_powo(multiple[i])
    # only if there is a match store result in data, otherwise put NA
    if(length(r$results)>0){
      # make dataframe
      tmp <- tidy(r)
      # add to list
      multiple_powo[[i]] <- tmp
    } else {
      multiple_powo[[i]] <- NA
    }
    write_rds(multiple_powo, "../Extra/Names/powo_multiple.R")
}} else {
  multiple_powo <- read_rds("../Extra/Names/powo_multiple.R")
}

# only get all with data
multiple_name <- multiple[!lengths(multiple_powo)<2]
multiple_powo_all <- multiple_powo
multiple_powo <- multiple_powo[!lengths(multiple_powo)<2]

# get the correct and accepted names
multiple_safe <- data.frame(old = character(), new= character(), accepted= character())
for(i in 1: length(multiple_powo)){
  dat <- multiple_powo[[i]]
  acc_name <- NA
  acc <- NA
  old <- NA
  if(nrow(dat) == 1){
    multiple_safe <- rbind(multiple_safe, data.frame(old= multiple_name[i], new = dat$name, accepted= dat$accepted))
  } else {
    acc_name <- dat$name[dat$accepted==T]
    acc_name <- ifelse(any(acc_name %in% multiple_name[i]), acc_name[match(acc_name, multiple_name[i])], acc_name)
    acc <- dat$accepted[dat$accepted==T]
    acc <- ifelse(length(acc)>1, any(dat$accepted[dat$name %in% acc_name]), acc)
    old <- multiple_name[i]
    multiple_safe <- rbind(multiple_safe, data.frame(old= old, new = acc_name, accepted= acc))
  }
}

names$powo[is.na(names$powo)] <- multiple_safe$new[match(names$waardplant[is.na(names$powo)], multiple_safe$old)]

# check which not yet
multiple_not <- multiple_safe[is.na(multiple_safe$new),]
# adapt --> all 18 checked
multiple_not$new <- multiple_not$old

# safe
names$powo[is.na(names$powo)] <- multiple_not$new[match(names$waardplant[is.na(names$powo)], multiple_not$old)]
```


Check remaining names
```{r}
# check and remove some names incorrectly added
lacking <- names[is.na(names$powo),]
lacking <- lacking[order(lacking$waardplant),]
lacking <- lacking[, c(1,2,5,6)]


lacking$powo <- c("Aconitum × cammarum","Aesculus x carnea","Allium christophii","Allium vineale","Eriocapitella × hybrida",
                  "Arctium minus","Arctium x ambiguum","Berberis × ottawensis","Brassica napus","Buddleja x weyeriana",
                  "Calamintha nepeta","Campylopus introflexus","Caryopteris x clandonensis","Centaurea cyanus",
                  "Citrus x aurantium","Clinopodium nepeta","Crocus x helenae","Crocus x luteus","Echinops banaticus","Erica x darleyensis",
                  "Erodium cicutarium","Fallopia x bohemica","Ficaria verna","Fragaria x ananassa","Gaillardia x grandiflora",
                  "Geranium x cantabrigiense","Geranium x magnificum","Geranium x oxonianum","Helianthus x laetiflorus",
                  "Heracleum sphondylium","Hyacinthoides x massartiana","Hyacinthoides x massartiana","Hypericum x desetangsii",
                  "Knautia arvensis","Lonicera fragrantissima","Medicago x varia","Mentha x verticillata",
                  "Mentha x piperita","Mentha x rotundifolia", "Narcissus pseudonarcissus",
                  "Nepeta x faassenii","Ononis spinosa","Osmanthus x burkwoodii","Petunia × atkinsiana", "Platanus x hispanica",
                  "Populus x canadensis","Poterium sanguisorba","Prunus avium","Prunus domestica",
                  "Prunus x fruticans","Rhododendron x praecox","Ribes x nidigrolaria","Rubus sect. Rubus","Salix babylonica",
                  "Salix cinerea",
                  "Salix x ambigua","Salix x friesiana","Salix x guinieri","Salix x reichardtii","Salix x rubra","Salix × pendulina nothof. tristis",
                  "Salix gmelinii","Salix gmelinii","Salvia x sylvestris",
                  "Solanum nigrum","Symphoricarpos x chenaultii","Symphyotrichum x versicolor","Symphytum grandiflorum",
                  "Symphytum x hidcotense","Symphytum x uplandicum","Taraxacum sect. Taraxacum","Teucrium x lucidrys",
                  "Tilia x europaea","Glandularia × hybrida","Veronica chamaedrys","Vicia sativa","Vicia tetrasperma" ,"Viola canina","Viola x wittrockiana"
                  ,"Vitis vinifera")


names$powo[is.na(names$powo)] <- lacking$powo[match(names$original_name[is.na(names$powo)], lacking$original_name)]
```

Again GBIF
```{r}
names$gbif <- gbif$species[match(names$waardplant, gbif$verbatim_name)]

# check differences
dif <- names[names$powo != names$gbif,]
dif <- dif[!is.na(dif$waardplant),]

names$powo[names$waardplant %in% dif$waardplant] <- dif$gbif[match(names$waardplant[names$waardplant %in% dif$waardplant],dif$waardplant)]
```


Genus
```{r}
# which have no genus or family
no_genus <- names[is.na(names$genus),]
no_family <- names[is.na(names$family),]

# first use gbif again
gbif <-  name_backbone_checklist(no_genus$powo)
no_genus$genus <- gbif$genus
no_genus$family <- gbif$family

# assign
names[is.na(names$genus),] <- no_genus

# no genus and family
no_genus <- names[is.na(names$genus) & is.na(names$family),]
no_genus$genus <- "Salix"
no_genus$family <- "Salicaceae"
names[is.na(names$genus) & is.na(names$family),] <- no_genus


# no genus
no_genus <- names[is.na(names$genus),]
vegdata <- vegdata::parse.taxa(no_genus$powo)
no_genus$genus <- vegdata$genus
names[is.na(names$genus),] <- no_genus


```


Check remaining names
```{r}
# some changes
names <- names[!names$powo =="Micrasterias decemdentata", ]
names$genus[names$waardplant == names$family] <- NA
names$genus[names$waardplant =="Bambusoideae"] <- NA
names[names$waardplant=="Lotus glaber", 6:7] <- data.frame("Lotus tenuis","Lotus")
names[names$waardplant=="Aronia", 7] <- data.frame("Aronia")
names[names$waardplant=="Stellaria", 7] <- data.frame("Stellaria")
names[names$waardplant=="Pastinaca", 7:6] <- data.frame("Pastinaca","Apiaceae")
names[names$waardplant=="Anchusa arvensis", 6:7] <- data.frame("Anchusa arvensis","Anchusa")
names[names$waardplant=="Trifolium", 7] <- data.frame("Trifolium")
names[names$waardplant=="Fallopia japonica + Fallopia sachalinensis + Fallopia x bohemica", 6:7] <- data.frame("Reynoutria × bohemica", "Reynoutria")
names[names$waardplant=="Eriobotrya japonica", 6:7] <- data.frame("Eriobotrya japonica", "Eriobotrya")
names[names$waardplant=="Perovskia ", 6] <- data.frame("Salvia")
names[names$waardplant=="Lychnis", 6] <- data.frame("Silene")
names[names$waardplant=="Silene armeria", 7] <- data.frame("Silene")
names[names$powo=="Salix alba", 7] <- data.frame("Salix")
names[names$powo=="Cnicus", 7] <- data.frame("Centaurea")
names[names$powo=="Convolvulus soldanella", 6:7] <- data.frame("Calystegia soldanella","Calystegia")
names[names$powo=="Inula conyzae", 6:7] <- data.frame("Pentanema squarrosum","Pentanema")
names[names$powo=="Leontodon saxatilis", 7] <- data.frame("Leontodon")
names[names$powo=="Peucedanum palustre", 7] <- data.frame("Peucedanum")
names[names$powo=="Potentilla rupestris", 6:7] <- data.frame("Drymocallis rupestris","Drymocallis")
names[names$powo=="Sedum sediforme", 6:7] <- data.frame("Petrosedum sediforme","Petrosedum")
names[names$powo=="Sedum rupestre", 6:7] <- data.frame("Petrosedum rupestre","Petrosedum")
names[names$powo=="Sorbus intermedia", 6:7] <- data.frame("Scandosorbus intermedia","Scandosorbus")
names[names$powo=="Brasica and Sinapis", 6] <- data.frame("Brassicaceae")
```


## Origin
```{r}
status <- read_csv("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/GLONAF/Belgium.csv", show_col_types = FALSE)
unique(status$status)
names$source <- NA
# assign Belgium
names$status <- status$status[match(names$powo, status$taxon_corrected)]
names$source[!is.na(names$status)] <- "GloNAF"
sum(!is.na(names$status))
```

```{r}
status <- readxl::read_excel("C:/Users/u0166342/Documents/Doctoraat/Impact/Extra data/Species names/Veronika_name_correction.xlsx")
extra <- status[status$Status.checked.VK =="extra",]
extra <- extra[!duplicated(extra$`Aggregated name`),]
status <- status[status$Region=="Belgium",]

# assign Belgium
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$powo[is.na(names$status)], status$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$gbif[is.na(names$status)] , status$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$original_name[is.na(names$status)] , status$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$powo[is.na(names$status)] , status$`Original EVA name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$gbif[is.na(names$status)] , status$`Original EVA name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$Status.checked.VK[match(names$original_name[is.na(names$status)] , status$`Original EVA name`)]
sum(!is.na(names$status))

# assign extra-European
names$status[is.na(names$status)] <- extra$Status.checked.VK[match(names$powo[is.na(names$status)] , extra$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$Status.checked.VK[match(names$original_name[is.na(names$status)] , extra$`Aggregated name`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$Status.checked.VK[match(names$powo[is.na(names$status)] , extra$`Original EVA name`)]
sum(!is.na(names$status))

names$source[!(is.na(names$status)) & is.na(names$source)] <- "Kalusova"
```


```{r}
status <- readxl::read_excel("C:/Users/u0166342/Documents/Doctoraat/Impact/Extra data/Species names/Irena_taxa_update.xlsx")
extra <- status[status$statusNew =="extra",]
extra <- extra[!duplicated(extra$species),]

status <- status[status$Region=="Belgium",]


# assign Belgium
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)], status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original_name[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original_name[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))

# assign extra-European
names$status[is.na(names$status)] <- extra$statusNew[match(names$powo[is.na(names$status)] , extra$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$statusNew[match(names$original_name[is.na(names$status)] , extra$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- extra$statusNew[match(names$powo[is.na(names$status)] , extra$`Matched concept`)]
sum(!is.na(names$status))

names$source[!(is.na(names$status)) & is.na(names$source)] <- "Kalusova"
```





```{r}
species_mab <- read.delim("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/Names/Aliens/taxon.txt")
status_mab <- read.delim("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/Names/Aliens/distribution.txt")

species_mab$species <- vegdata::taxname.removeAuthors(species_mab$scientificName)

status<- species_mab
status$statusNew <- "Alien"

# assign Belgium
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)], status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original_name[is.na(names$status)] , status$species)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$powo[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$gbif[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))
names$status[is.na(names$status)] <- status$statusNew[match(names$original_name[is.na(names$status)] , status$`Matched concept`)]
sum(!is.na(names$status))

names$source[!(is.na(names$status)) & is.na(names$source)] <- "Manual alien plants Belgium"

```



```{r}
names$status[names$status %in% c("Naturalized","Invasive","intra","extra")] <- "Alien"
names$status[names$status %in% c("native")] <- "Native"


check <- names[!is.na(names$status),]


# here most of my assignments seem correct, so keep these
control <- check[check$status_estelle != check$status,]

# remaining
remaining <-names[is.na(names$status),]
remaining_aliens <- remaining[remaining$status_estelle =="Alien",]
# I checked some, all were indeed alien

# Assign
names$source[is.na(names$source)] <- "waarnemingen"
names$status[is.na(names$status)] <- names$status_estelle[is.na(names$status)]

names <- names[, -c(9)]
names <- names[, -c(5)]
colnames(names)[5] <- 'name_willem'
colnames(names)[8] <- 'status_willem'
```

```{r}
status <- read_csv("C:/Users/u0166342/Documents/Doctoraat/Observation/Extra/GLONAF/Belgium.csv", show_col_types = FALSE)
status <- status[status$status=="Invasive",]

names$invasive_willem <- status$status[match(names$name_willem, status$taxon_corrected)]

# based on the invasive alien species in Belgium document
species_vector_invasives <- c(
  "Acer negundo",
  "Acer rufinerve",
  "Ailanthus altissima",
  "Akebia quinata",
  "Ambrosia artemisiifolia",
  "Amelanchier lamarckii",
  "Aster lanceolatus",
  "Aster novi-belgii",
  "Aster x salignus",
  "Azolla filiculoides",
  "Baccharis halimifolia",
  "Bidens frondosa",
  "Buddleja davidii",
  "Cabomba caroliniana",
  "Carpobrotus spp.",
  "Cornus sericea",
  "Cotoneaster horizontalis",
  "Crassula helmsii",
  "Cyperus eragrostis",
  "Duchesnea indica",
  "Echinocystis lobata",
  "Egeria densa",
  "Elaeagnus angustifolia",
  "Elodea canadensis",
  "Elodea nuttallii",
  "Epilobium ciliatum",
  "Fallopia japonica",
  "Fallopia sachalinensis",
  "Fallopia x bohemica",
  "Fraxinus pennsylvanica",
  "Gaillardia x grandiflora",
  "Helianthus tuberosus",
  "Heracleum mantegazzianum",
  "Hyacinthoides hispanica",
  "Hydrocotyle ranunculoides",
  "Impatiens glandulifera",
  "Impatiens parviflora",
  "Koenigia polystachya",
  "Lagarosiphon major",
  "Lemna minuta",
  "Lonicera japonica",
  "Ludwigia grandiflora",
  "Ludwigia peploides",
  "Lupinus polyphyllus",
  "Lysichiton americanus",
  "Mahonia aquifolium",
  "Mimulus guttatus",
  "Myriophyllum aquaticum",
  "Myriophyllum heterophyllum",
  "Oenothera spp.",
  "Parthenocissus spp.",
  "Phytolacca americana",
  "Prunus laurocerasus",
  "Prunus serotina",
  "Quercus rubra",
  "Rhododendron ponticum",
  "Rhus typhina",
  "Robinia pseudoacacia",
  "Rosa rugosa",
  "Rudbeckia laciniata",
  "Senecio inaequidens",
  "Solidago canadensis",
  "Solidago gigantea",
  "Spiraea alba",
  "Spiraea douglasii",
  "Spiraea tomentosa",
  "Spiraea x billardii"
)

names$invasive_willem[is.na(names$invasive_willem) & names$name_willem %in% species_vector_invasives] <- "Invasive"

sum(!is.na(names$invasive_willem))

write_csv(names, "../Extra/Names/Estelle_origins.csv")
```


