---
title: "Network analysis"
author: "Willem"
date: "2025-06-18"
output: html_document
---
# 1 START
This is a preliminary analysis, before analysing associated plant species via plantnet

Load
```{r}
library(tidyverse)
library(bipartite)
```

Remove all available data
```{r}
rm(list=ls())
```

Load in data
```{r}
polplant <- read_csv("../Data/polplant.csv")
fast=F
if(fast){
  polplant <- polplant[runif(nrow(polplant)) >0.99,]
}

# create interaction matrix
interaction_matrix <- polplant |>dplyr::count(powo, gbif) |> tidyr::pivot_wider(names_from = gbif,values_from = n,values_fill = list(n = 0))
interaction_matrix <- as.data.frame(interaction_matrix)
rownames(interaction_matrix) <- interaction_matrix$powo
interaction_matrix <- interaction_matrix[, -1]
```


# 2 Network indices
Which chose?
- connectance (depend on specialisation but also sampling intensity)
- degree also 
- weighted vs binary (use less information, but not necessarily more robust, but more affected by sampling bias)

Table
- connectance
- links per species
- nestedness
- mean number of partners
- degree

Network level
```{r}
# COMPARTMENTS
# compartments are sub sets not connected to another one
comp_number<- networklevel(interaction_matrix, index= "number of compartments")
# visualise using visweb
# visweb(interaction_matrix)
# shannon diversity of compartment sizes
#comp_div<- networklevel(interaction_matrix, index= "compartment diversity")

# NESTEDNESS
# 0 highly nested, 1 pure chaos
nestedness<- networklevel(interaction_matrix, index= "nestedness")
# for null models, use nested
null_model <- nested(interaction_matrix)
# higher values more nested here
#nestedness_NODF <- networklevel(interaction_matrix, index= "NODF")
# interaction frequencies accounted for (0 chaos to 1 nested)
#nestedness_weighted<- networklevel(interaction_matrix, index= "weighted nestedness")
# weighted NODF (but takes long to calculate)
#nestedness_NODF_weighted<- networklevel(interaction_matrix, index= "weighted NODF")
# other options are spectral radius

# SPECIALISATION AND DIVERSITY
# dependence assymetry --> specialisation 
# alternative ISAmethod= 'Bluethgen' --> positive = higher dependence in higher trophic 
# more info specieslevel --> index push/pull affecting of being affected
ISA<- networklevel(interaction_matrix, index= "ISA", ISAmethod= 'Bluethgen')
# assymetry (high vs low) of specialisation
# Bluethgen --> weight by abundance or log -_> log transformation d values
# positive --> higher specialisation higher trophic level
SA <- networklevel(interaction_matrix, index= "SA", SAmethod= 'Bluethgen')
# balance between numbers in level, negative more lower trophic (plants here) species
#web_assymetry<- networklevel(interaction_matrix, index= "web asymmetry")
#int_eveness  <- networklevel(interaction_matrix, index= "interaction evenness")

# CONNECTANCE 
# realised proportion of possible links
connectance <- networklevel(interaction_matrix, index= "connectance")
# average of vulnerability and generality
#link_dens <- networklevel(interaction_matrix, index= "linkage density")
# lingage density / number species
#weight_connect<- networklevel(interaction_matrix, index= "weighted connectance")
# number realised links divided by number possible (all, higher, lower)
cluster_coef <-  networklevel(interaction_matrix, index= "cluster coefficient")
#mean number of links per species
#links_pp<- networklevel(interaction_matrix, index= "links per species")
```

Group level
```{r}
# BASIC
num_sp <- networklevel(interaction_matrix, index= "number of species")
#weight_clust_coef <- networklevel(interaction_matrix, index= "weighted cluster coefficient")

# generalitivity --> mean effective number of LL species per HL species, HL per LL species for vulenerability
gen <-networklevel(interaction_matrix, index= "generality")
# area below secondary extinction sequency
robustness <- networklevel(interaction_matrix, index= "robustness")
# mean number of co-occuoancies
#togetherness<- networklevel(interaction_matrix, index= "togetherness")
# niche overlap --> mean similarity
niche_overlap<- networklevel(interaction_matrix, index= "niche overlap")
```

Species level
```{r}
# number of links per species
degree <- specieslevel(interaction_matrix, "degree")
# sum of dependencies
strength <-  specieslevel(interaction_matrix, "species strength")

# multiple of lower trophic 
specieslevel(interaction_matrix, level="lower", index=c("normalised degree", "PDI", "effective partners"), PDI.normalise=F)
```



Modules
```{r}
mod <- computeModules(interaction_matrix)
plotModuleWeb(mod)
```

