---
title: "Network analysis"
author: "Willem"
date: "2025-06-18"
output: html_document
---
# START
This is a preliminary analysis, before analysing associated plant species via plantnet

Load
```{r}
library(tidyverse)
library(ggpubr)
library(bipartite)
library(ggridges)
library(multcomp)
library(lme4)
library(lmerTest)
library(DHARMa)
library(glmmTMB)
library(iNEXT)
library(emmeans)
library(sf)
library(sp)
library(rnetcarto)
library(ggpubr)
library(ggExtra)
```

Remove all available data
```{r}
rm(list=ls())
```

Load in data
```{r}
# load data
interactions_all <- read_csv("../Results/Interaction_all.csv", show_col_types = FALSE)
interactions <- interactions_all

# remove old data
interactions <- interactions[!interactions$date < "2010-01-01", ]

# remove singletons
interactions <- interactions |>   group_by(name, plant_taxa, eco_region) |>  filter(n() >= 2) |>  ungroup()

# subdivide data 
fast =F
if(fast){
  interactions <- interactions[runif(nrow(interactions)) > 0.90,]
}


# get specific data only
interactions <- interactions |> 
  dplyr::select(id, eco_region, date, name, plant_taxa, euromed_region, plant_genus, plant_family,plant_rank_final, pol_genus, pol_family, 
                pol_rank, lat, lng)
interactions |> group_by(eco_region, euromed_region) |> summarise(n=n())
```


```{r}
# assign native status
status <- read_csv("../Extra/Names/Interaction_alien_status.csv",show_col_types = FALSE)
status <- status |>  distinct(plant_taxa, euromed_region, .keep_all = TRUE)
interactions <- left_join(  interactions, status[, c("plant_taxa", "euromed_region", "statusNew")], by = c("plant_taxa", "euromed_region"))
colnames(interactions)[15] <- "status"

# check numbers
table(interactions$status)
table(status$statusNew)

# check overall percentage
sum(!is.na(interactions$status))/nrow(interactions)
sum(!is.na(status$statusNew))/nrow(status)

# only keep those with species classified
interactions <- interactions[!is.na(interactions$status),]

test <- interactions |> distinct(plant_taxa, status)
table(test$status)
```


```{r}
# only keep species level observations for now
interactions <- interactions[interactions$plant_rank_final=="Species" & interactions$pol_rank== "Species",]

# create interaction matrix
# depends on what you want

# all subdivisions
interaction_matrix_all<- interactions|>dplyr::count(name, plant_taxa,eco_region, euromed_region, status) |> 
  tidyr::pivot_wider(names_from = name,values_from = n, values_fill = list(n = 0))
interaction_matrix <- as.data.frame(interaction_matrix_all)

# only status subdivisions
interaction_matrix_status <- interactions |> dplyr::count(name, plant_taxa, status) |> 
  tidyr::pivot_wider(names_from = name,values_from = n, values_fill = list(n = 0))
interaction_matrix <- as.data.frame(interaction_matrix_status)

table(interaction_matrix$status)
```

# 1 SPECIES 
## 1.1 Calculate relevant metrics
```{r}
# calculate all relevant metrics
mat <- interaction_matrix[, -c(1:2)]
nL <- nrow(mat)
nH <- ncol(mat)


deg <- specieslevel(mat,"degree")
ND <- specieslevel(mat,"normalised degree")
strg <- specieslevel(mat,"species strength")

NSI <- specieslevel(mat,"NSI")
BC <- specieslevel(as.matrix(mat),"betweenness")
RR <- specieslevel((mat),"resource range")

number_obs_plant <- interactions |> dplyr::group_by(plant_taxa) |> summarise(n=n())
number_obs_pol <- interactions |> dplyr::group_by(name) |> summarise(n=n())

# assign
higher <- data.frame(species=colnames(mat), n=nH,degree=deg$`higher level`,
                     ND=ND$`higher level`,
                     strength=strg$`higher level`, NSI=NSI$`higher level`,  
                     RR = RR$`higher level`,
                     BC = BC$`higher level`, 
                     number = number_obs_plant$n[match(colnames(mat),
                                                       number_obs_pol$name)])
lower <- data.frame(species= interaction_matrix$plant_taxa, n=nL, degree=deg$`lower level`,
                    ND=ND$`lower level`, strength=strg$`lower level`, NSI=NSI$`lower level`,
                    RR = RR$`lower level`, BC = BC$`lower level`,
                    number = number_obs_plant$n[match(interaction_matrix$plant_taxa, number_obs_plant$plant_taxa)])

# assign status
lower$status <- interaction_matrix$status
# assign genera and families
lower$genus <- interactions$plant_genus[match(lower$species, interactions$plant_taxa)]
lower$family <- interactions$plant_family[match(lower$species, interactions$plant_taxa)]

lower$node.specialisation.index.NSI[lower$node.specialisation.index.NSI=="NaN"]<- NA
# check correlations
cor(lower[!is.na(lower$node.specialisation.index.NSI), 3:10], method="spearman")
cor(lower[!is.na(lower$node.specialisation.index.NSI), 3:10], method="pearson")
```


## 1.2 Degree
```{r}
#  check role degree
sign_status <- glmer(degree ~ status + (1|family/genus ), lower, family = poisson(link= "log"))
sign_status <- glmmTMB(degree ~ status + (1|family/genus ), lower, family = nbinom2())
# should we add number of observations
#sign_status <- glmmTMB(degree ~ status +  log10(number) + (1|family/genus ), lower, family = nbinom2())
summary(sign_status)
plot(simulateResiduals(sign_status))

# check significance
emm <- emmeans(sign_status, pairwise ~ status)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
cld_results

count <- data.frame(table(lower$status))
count$let <- cld_results$.group[match(count$Var1, cld_results$status)]

degree_status <- ggplot(lower, aes(x= status, y= degree, col = status, fill = status))+
  geom_violin(scale= "area")+
  ggpubr::theme_pubr()+
  geom_boxplot(width= 0.2, alpha=0.8, fill="white")+
  scale_colour_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native")) +
  scale_fill_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native"))+
  stat_summary(fun= "mean", geom = "point", aes(group= status), size=2)+
  theme(legend.position = "none")+
  ylab("Degree")+
  xlab(NULL)+
  scale_x_discrete(labels=c("Extra-\nEuropean", "Intra-\nEuropean", "Native"))+
  geom_text(data = count, aes(x = Var1, y = max(lower$degree)*1.05, label = let), 
            inherit.aes = FALSE, hjust = 0.5) +
  geom_text(data = count, aes(x = Var1, y = max(lower$degree)*1.15, label = paste("n= ", Freq, sep="")), 
            inherit.aes = FALSE, hjust = 0.5) 

ggsave("../Images/Overall_status_degree.png", degree_status, width= 5, height =3)
```


## 1.3 Resource range
```{r}
#  check role degree
sign_status <- glmmTMB(resource.range ~ status + (1|family/genus ), lower, family = ordbeta())
summary(sign_status)
plot(simulateResiduals(sign_status))

# check significance
emm <- emmeans(sign_status, pairwise ~ status)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
cld_results

count <- data.frame(table(lower$status))
count$let <- cld_results$.group[match(count$Var1, cld_results$status)]

resource_range_status <- ggplot(lower, aes(x= status, y= resource.range, col = status, fill = status))+
  geom_violin(scale= "area")+
  ggpubr::theme_pubr()+
  geom_boxplot(width= 0.2, alpha=0.8, fill="white")+
  scale_colour_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native")) +
  scale_fill_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native"))+
  stat_summary(fun= "mean", geom = "point", aes(group= status), size=2)+
  theme(legend.position = "none")+
  ylab("Resource range")+
  xlab(NULL)+
  scale_x_discrete(labels=c("Extra-\nEuropean", "Intra-\nEuropean", "Native"))+
  #geom_text(data = count, aes(x = Var1, y = max(lower$resource.range)*1.01, label = let), 
  #          inherit.aes = FALSE, hjust = 0.5) +
  geom_text(data = count,aes(x=Var1, y = max(lower$resource.range)*1.03,label =paste("n= ",Freq, sep="")), 
            inherit.aes = FALSE, hjust = 0.5) 

ggsave("../Images/Overall_status_resource_range.png", resource_range_status, width= 5, height =3)
```


## 1.3 BC
```{r}
#  check role degree
sign_status <- glmmTMB(BC.betweenness ~ status + (1|family/genus ), lower, family = ordbeta())
summary(sign_status)
plot(simulateResiduals(sign_status))

# check significance
emm <- emmeans(sign_status, pairwise ~ status)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
#cld_results

count <- data.frame(table(lower$status))
count$let <- cld_results$.group[match(count$Var1, cld_results$status)]

BC_status <- ggplot(lower, aes(x= status, y= BC.betweenness, col = status, fill = status))+
  geom_violin(scale= "area")+
  ggpubr::theme_pubr()+
  geom_boxplot(width= 0.2, alpha=0.8, fill="white")+
  scale_colour_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native")) +
  scale_fill_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native"))+
  stat_summary(fun= "mean", geom = "point", aes(group= status), size=2)+
  theme(legend.position = "none")+
  ylab("Betweenness\ncentrality")+
  xlab(NULL)+
  scale_x_discrete(labels=c("Extra-\nEuropean", "Intra-\nEuropean", "Native"))+
  #geom_text(data = count, aes(x = Var1, y = max(lower$BC.betweenness)*1.01, label = let), 
  #          inherit.aes = FALSE, hjust = 0.5) +
  geom_text(data = count,aes(x=Var1, y = max(lower$BC.betweenness)*1.05,label =paste("n= ",Freq, sep="")), 
            inherit.aes = FALSE, hjust = 0.5) 

ggsave("../Images/Overall_status_BC.png", BC_status, width= 5, height =3)
```



## 1.4 NSI
```{r}
#  check role degree
sign_status <- glmmTMB(node.specialisation.index.NSI-1 ~ status + (1|family/genus ), lower, 
                       family = Gamma(link="log"))
summary(sign_status)
plot(simulateResiduals(sign_status))

# check significance
emm <- emmeans(sign_status, pairwise ~ status)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
#cld_results

count <- data.frame(table(lower$status))
count$let <- cld_results$.group[match(count$Var1, cld_results$status)]

NSI_status <- ggplot(lower, aes(x= status, y= node.specialisation.index.NSI , col = status, fill = status))+
  geom_violin(scale= "area")+
  ggpubr::theme_pubr()+
  geom_boxplot(width= 0.2, alpha=0.8, fill="white")+
  scale_colour_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native")) +
  scale_fill_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native"))+
  stat_summary(fun= "mean", geom = "point", aes(group= status), size=2)+
  theme(legend.position = "none")+
  ylab("Node Specialisation\nIndex")+
  xlab(NULL)+
  scale_x_discrete(labels=c("Extra-\nEuropean", "Intra-\nEuropean", "Native"))+
  geom_text(data = count, aes(x = Var1, y = max(lower$node.specialisation.index.NSI, na.rm =T)*1.05, 
                              label = let), inherit.aes = FALSE, hjust = 0.5) +
  geom_text(data = count,aes(x=Var1, y = max(lower$node.specialisation.index.NSI, na.rm=T)*1.10,
                             label =paste("n= ",Freq, sep="")), inherit.aes = FALSE, hjust = 0.5) 

ggsave("../Images/Overall_status_NSI.png", NSI_status, width= 5, height =3)
```


## 1.5 Strength
```{r}
#  check role degree
sign_status <- glmmTMB(log(species.strength)~ status + (1|family/genus ), lower, family = gaussian())
summary(sign_status)
plot(simulateResiduals(sign_status))

# check significance
emm <- emmeans(sign_status, pairwise ~ status)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
#cld_results

count <- data.frame(table(lower$status))
count$let <- cld_results$.group[match(count$Var1, cld_results$status)]

strength_status <- ggplot(lower, aes(x= status, y= log(species.strength) , col = status, fill = status))+
  geom_violin(scale= "area")+
  ggpubr::theme_pubr()+
  geom_boxplot(width= 0.2, alpha=0.8, fill="white")+
  scale_colour_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native")) +
  scale_fill_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native"))+
  stat_summary(fun= "mean", geom = "point", aes(group= status), size=2)+
  theme(legend.position = "none")+
  ylab("Log species strength")+
  xlab(NULL)+
  scale_x_discrete(labels=c("Extra-\nEuropean", "Intra-\nEuropean", "Native"))+
  geom_text(data = count, aes(x = Var1, y = max(log(lower$species.strength), na.rm =T)*1.15, 
                              label = let), inherit.aes = FALSE, hjust = 0.5) +
  geom_text(data = count,aes(x=Var1, y = max(log(lower$species.strength), na.rm=T)*1.50,
                             label =paste("n= ",Freq, sep="")), inherit.aes = FALSE, hjust = 0.5) 

ggsave("../Images/Overall_status_strength.png", strength_status, width= 5, height =3)
```



```{r}
# combine
p <- ggpubr::ggarrange(degree_status, strength_status, BC_status, NSI_status, ncol= 2, nrow= 2, 
               labels= letters[1:4], align= "hv")

ggsave("../Images/Overall_status.svg", p, width= 7, height =5)
```


# 2 RE
## 2.1 ALL
```{r}
interact_long <- interactions[,c(4:5,15)]
interact_long <- interact_long |> group_by(plant_taxa, name, status) |> summarise(n=n(), .groups = "drop")

interact_list_status <- list(native = interact_long |> filter(status == "native") |> pull(n),
                             intra  = interact_long |> filter(status == "intra")  |> pull(n),
                             extra  = interact_long |> filter(status == "extra")  |> pull(n))
view(interact_list_status$intra, n=1500)
RE <- iNEXT(interact_list_status, q =0, datatype = "abundance" )
#write_rds(RE, "../Extra/Intermediate/Rarefaction_accumulation_overall_status.Rdata")
RE <- read_rds("../Extra/Intermediate/Rarefaction_accumulation_overall_status.Rdata")

cols <- c("extra"  = "#F07167","intra"  = "#7CAE00","native" = "#00A9A5")
RE_int <- ggiNEXT(RE)+
  ggpubr::theme_pubr()+
  xlab("Number of observed interactions")+
  scale_color_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native")) +
  scale_fill_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native"))+
  scale_shape(labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native"))+
  scale_x_continuous(limits= c(0, 4.5*10^5),
                     breaks = c(0,1*10^5, 2*10^5, 3*10^5, 4*10^5),
                     labels= c("0",expression(1 %*% 10^5),expression(2 %*% 10^5),expression(3 %*% 10^5), expression(4 %*% 10^5)))+
  ylab("Unique interactions")+
  theme(legend.position = "right")

ggsave("../Images/Rarefaction_overall_status.png", RE_int, width= 6, height=3)

```

```{r}
interact_long <- interactions[,c(5,15)]
interact_long <- interact_long |> group_by(plant_taxa, status) |> summarise(n=n(), .groups = "drop")

interact_list_status <- list(native = interact_long |> filter(status == "native") |> pull(n),
                             intra  = interact_long |> filter(status == "intra")  |> pull(n),
                             extra  = interact_long |> filter(status == "extra")  |> pull(n))

RE <- iNEXT(interact_list_status, q =0, datatype = "abundance" )
#write_rds(RE, "../Extra/Intermediate/Rarefaction_accumulation_plant_status_plant.Rdata")
RE <- read_rds("../Extra/Intermediate/Rarefaction_accumulation_plant_status_plant.Rdata")
cols <- c("extra"  = "#F07167","intra"  = "#7CAE00","native" = "#00A9A5")
RE_plant <- ggiNEXT(RE)+
  ggpubr::theme_pubr()+
  xlab("Number of observed interactions")+
  scale_color_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native")) +
  scale_fill_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native"))+
  scale_shape(labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native"))+
  scale_x_continuous(limits= c(0, 4.5*10^5),
                     breaks = c(0,1*10^5, 2*10^5, 3*10^5, 4*10^5),
                     labels= c("0",expression(1%*%10^5),expression(2%*%10^5),expression(3%*%10^5), expression(4%*%10^5)))+
  ylab("Unique plant species")+
  theme(legend.position = "right")

ggsave("../Images/Rarefaction_overall_status_plant.png", RE_plant, width= 6, height=3)
```

```{r}
interact_long <- interactions[,c(4,15)]
interact_long <- interact_long |> group_by(name) |> summarise(n=n(), .groups = "drop")

interact_list_status <- list(pol = interact_long  |> pull(n))

RE <- iNEXT(interact_list_status, q =0, datatype = "abundance" )
#write_rds(RE, "../Extra/Intermediate/Rarefaction_accumulation_overall_status_bee.Rdata")
RE <- read_rds("../Extra/Intermediate/Rarefaction_accumulation_overall_status_bee.Rdata")

RE_bee <- ggiNEXT(RE)+
  ggpubr::theme_pubr()+
  xlab("Number of observed interactions")+
  scale_x_continuous(limits= c(0, 4.5*10^5),
                     breaks = c(0,1*10^5, 2*10^5, 3*10^5, 4*10^5),
                     labels= c("0",expression(1 %*% 10^5),expression(2 %*% 10^5),expression(3 %*% 10^5), expression(4 %*% 10^5)))+
  ylab("Unique bee species")+
  theme(legend.position = "none")+
  guides( colour = "none",shape  = "none",fill= "none")

ggsave("../Images/Rarefaction_overall_status_bee.png", RE_bee, width= 6, height=3)
```


```{r}
RE_plant <- RE_plant + 
  theme(axis.title.x = element_blank())+ 
  scale_x_continuous(limits= c(0, 4.5*10^5),
                     breaks = c(0, 2*10^5, 4*10^5),
                     labels= c("0",expression(2%*%10^5), expression(4%*%10^5)))
RE_bee   <- RE_bee   + theme(axis.title.x = element_blank())+ 
  scale_x_continuous(limits= c(0, 4.5*10^5),
                     breaks = c(0, 2*10^5, 4*10^5),
                     labels= c("0",expression(2%*%10^5), expression(4%*%10^5)))
RE_int   <- RE_int   + theme(axis.title.x = element_blank())+ 
  scale_x_continuous(limits= c(0, 4.5*10^5),
                     breaks = c(0, 2*10^5, 4*10^5),
                     labels= c("0",expression(2%*%10^5), expression(4%*%10^5)))

p <- ggarrange(RE_plant, RE_bee, RE_int, ncol=3, nrow=1, labels = letters[1:3], align= "hv", common.legend = T,
               legend ="top")
p <- annotate_figure( p, bottom = text_grob("Number of observed interactions"))
p
ggsave("../Images/Rarefaction_overall_status_all.png", p, width= 8, height=4)
```


## 2.2 Grid
```{r}
# make regions
medRegions <- read_sf("~/Doctoraat/Impact/Extra data/EURO+MED/Regions", "Emed_regions")
medRegions <- st_transform(medRegions, CRS("+proj=longlat +datum=WGS84"))

# make coordinates
coord <- interactions |> dplyr::select("name","plant_taxa", "status","lat", "lng")

# make geometrical dataset
coord <- st_as_sf(coord, coords = c("lng","lat"),crs = 4326)
grid <- st_make_grid(medRegions, cellsize = 1, square = TRUE) %>% st_sf()
grid$name_grid <- 1:nrow(grid)

# group
grid_counts <- st_join(coord, grid, join = st_intersects) 
```

```{r}
# make interaction column
grid_counts <- grid_counts |>  mutate(interaction = paste(plant_taxa, name, sep = "_"))
# group per region and status level
grid_counts_int <- grid_counts |> group_by(interaction, status, name_grid) |> summarise(n=n())
# remove geometry, necessary to calculate later
grid_counts_int <- grid_counts_int |> st_drop_geometry()

# filter and make list of matrices
grid_list_status <- list(native = grid_counts_int |> filter(status == "native") |> 
                           tidyr::pivot_wider(names_from = interaction, values_from = n,
                                              values_fill = 0),
                         intra  = grid_counts_int |> filter(status == "intra")  |> 
                           tidyr::pivot_wider(names_from = interaction, values_from = n,
                                              values_fill = 0),
                         extra  = grid_counts_int |> filter(status == "extra")  |> 
                           tidyr::pivot_wider(names_from = interaction, values_from = n,
                                              values_fill = 0))

# remove unneccessary columns
grid_list_status_mat <- list(native = (grid_list_status[[1]][,-c(1:2)]),
                         intra = (grid_list_status[[2]][,-c(1:2)]),
                         extra = (grid_list_status[[3]][,-c(1:2)]))

# make binary
grid_list_status_mat_binary <- lapply(
  grid_list_status_mat, function(x) { 
    x[x > 0] <- 1
    x
  }
)

# transpose datafram
grid_list_status_mat_binary <- list(native = as.data.frame(t(grid_list_status_mat_binary$native)),
                                    intra = as.data.frame(t(grid_list_status_mat_binary$intra)),
                                    extra = as.data.frame(t(grid_list_status_mat_binary$extra)))


# calculate RE
RE <-  iNEXT((grid_list_status_mat_binary), q =0, datatype = "incidence_raw" )
#write_rds(RE, "../Extra/Intermediate/Rarefaction_accumulation_grid_status.Rdata")
RE <- read_rds("../Extra/Intermediate/Rarefaction_accumulation_grid_status.Rdata")

# assign colours again
cols <- c("extra"  = "#F07167","intra"  = "#7CAE00","native" = "#00A9A5")

# calculate
RE_int_grid <- ggiNEXT(RE)+
  ggpubr::theme_pubr()+
  xlab("Number of grid cells")+
  scale_color_manual(values = cols, labels = c(extra  = "Extra-European",
                                               intra  = "Intra-European", native = "Native")) +
  scale_fill_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European",
                                              native = "Native"))+
  scale_shape(labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native"))+
  ylab("Unique interactions")+
  theme(legend.position = "right")

# save
ggsave("../Images/Rarefaction_grid_status.png", RE_int_grid, width= 6, height=3)
```


```{r}
# create interaction accumulation curve
sp_native <- specaccum(as.data.frame(grid_list_status_mat$native), method = "random")
sp_intra <- specaccum(as.data.frame(grid_list_status_mat$intra), method = "random")
sp_extra <- specaccum(as.data.frame(grid_list_status_mat$extra), method = "random")

# make it a data frame for plotting
accum_to_df <- function(x, status){data.frame(sites = x$sites,richness = x$richness,sd = x$sd,status = status )}
# apply function
df_accum <- bind_rows(accum_to_df(sp_native, "native"), accum_to_df(sp_intra,  "intra"),accum_to_df(sp_extra,  "extra"))

# get accumulation curve
ACC_all <- ggplot(df_accum, aes(x = sites, y = richness, colour = status, fill = status)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = richness - sd,
                  ymax = richness + sd),
              alpha = 0.2,
              colour = NA) +
  ggpubr::theme_pubr() +
  labs(x = "Number of grid cells",
       y = "Accumulated unique interactions")+
  scale_color_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native")) +
  scale_fill_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native"))+
  theme(legend.position = "right")
```


```{r}
# group per region and status level
grid_counts_int <- grid_counts |> group_by(plant_taxa, status, name_grid) |> summarise(n=n())
# remove geometry, necessary to calculate later
grid_counts_int <- grid_counts_int |> st_drop_geometry()

# filter and make list of matrices
grid_list_status <- list(native = grid_counts_int |> filter(status == "native") |> 
                           tidyr::pivot_wider(names_from = plant_taxa, values_from = n,
                                              values_fill = 0),
                         intra  = grid_counts_int |> filter(status == "intra")  |> 
                           tidyr::pivot_wider(names_from = plant_taxa, values_from = n,
                                              values_fill = 0),
                         extra  = grid_counts_int |> filter(status == "extra")  |> 
                           tidyr::pivot_wider(names_from = plant_taxa, values_from = n,
                                              values_fill = 0))

# remove unneccessary columns
grid_list_status_mat <- list(native = (grid_list_status[[1]][,-c(1:2)]),
                         intra = (grid_list_status[[2]][,-c(1:2)]),
                         extra = (grid_list_status[[3]][,-c(1:2)]))

# make binary
grid_list_status_mat_binary <- lapply(
  grid_list_status_mat, function(x) { 
    x[x > 0] <- 1
    x
  }
)

# transpose datafram
grid_list_status_mat_binary <- list(native = as.data.frame(t(grid_list_status_mat_binary$native)),
                                    intra = as.data.frame(t(grid_list_status_mat_binary$intra)),
                                    extra = as.data.frame(t(grid_list_status_mat_binary$extra)))


# calculate RE
RE <-  iNEXT((grid_list_status_mat_binary), q =0, datatype = "incidence_raw" )
#write_rds(RE, "../Extra/Intermediate/Rarefaction_accumulation_grid_status_plant.Rdata")
RE <- read_rds("../Extra/Intermediate/Rarefaction_accumulation_grid_status_plant.Rdata")


# assign colours again
cols <- c("extra"  = "#F07167","intra"  = "#7CAE00","native" = "#00A9A5")

# calculate
RE_plant_grid <- ggiNEXT(RE)+
  ggpubr::theme_pubr()+
  xlab("Number of grid cells")+
  scale_color_manual(values = cols, labels = c(extra  = "Extra-European",
                                               intra  = "Intra-European", native = "Native")) +
  scale_fill_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European",
                                              native = "Native"))+
  scale_shape(labels = c(extra  = "Extra-European",intra  = "Intra-European", native = "Native"))+
  ylab("Unique plant species")+
  theme(legend.position = "right")

# save
ggsave("../Images/Rarefaction_grid_status_plant.png", RE_plant_grid, width= 6, height=3)
```


```{r}
# group per region and status level
grid_counts_int <- grid_counts |> group_by(name, name_grid) |> summarise(n=n())
# remove geometry, necessary to calculate later
grid_counts_int <- grid_counts_int |> st_drop_geometry()

# filter and make list of matrices
grid_list_status <- list(native = grid_counts_int |>
                           tidyr::pivot_wider(names_from = name, values_from = n,
                                              values_fill = 0))

# remove unneccessary columns
grid_list_status_mat <- list(native = (grid_list_status[[1]][,-c(1)]))

# make binary
grid_list_status_mat_binary <- lapply(
  grid_list_status_mat, function(x) { 
    x[x > 0] <- 1
    x
  }
)

# transpose datafram
grid_list_status_mat_binary <- list(native = as.data.frame(t(grid_list_status_mat_binary$native)))


# calculate RE
RE <-  iNEXT((grid_list_status_mat_binary), q =0, datatype = "incidence_raw" )
#write_rds(RE, "../Extra/Intermediate/Rarefaction_accumulation_grid_status_bee.Rdata")
RE <- read_rds("../Extra/Intermediate/Rarefaction_accumulation_grid_status_bee.Rdata")


# calculate
RE_bee_grid <- ggiNEXT(RE)+
  ggpubr::theme_pubr()+
  xlab("Number of grid cells")+
  ylab("Unique bee species")+
  theme(legend.position = "right")+
  guides( colour = "none",shape  = "none",fill= "none")

# save
ggsave("../Images/Rarefaction_grid_status_bee.png", RE_int_grid, width= 6, height=3)
```


```{r}
RE_plant_grid <- RE_plant_grid + 
  theme(axis.title.x = element_blank())+ 
  scale_x_continuous(breaks = c(0, 500, 1000))
RE_bee_grid   <- RE_bee_grid   + 
  theme(axis.title.x = element_blank())+ 
  scale_x_continuous(breaks = c(0, 500, 1000))
RE_int_grid   <- RE_int_grid   + 
  theme(axis.title.x = element_blank())+ 
  scale_x_continuous(breaks = c(0, 500, 1000))

p <- ggpubr::ggarrange(RE_plant_grid, RE_bee_grid, RE_int_grid, ncol=3, nrow=1, labels = letters[1:3], align= "hv", common.legend = T,
               legend ="top")
p <- ggpubr::annotate_figure( p, bottom = text_grob("Number of grid cells"))
p
ggsave("../Images/Rarefaction_grid_status_all.png", p, width= 8, height=4)
```



## 2.3 Accumulate
```{r}
# make interaction column
interactions_long <- interactions |> dplyr::select(id, plant_taxa, name, status)
interactions_long <- interactions_long |>  mutate(interaction = paste(plant_taxa, name, sep = "_"))
interactions_long <- interactions_long |>  group_by(id, interaction, status, ) |>  summarise(n=n())

# filter and make list of matrices
grid_list_status <- list(native = interactions_long |> filter(status == "native") |> 
                           tidyr::pivot_wider(names_from = plant_taxa, values_from = n,
                                              values_fill = 0),
                         intra  = interactions_long |> filter(status == "intra")  |> 
                           tidyr::pivot_wider(names_from = plant_taxa, values_from = n,
                                              values_fill = 0),
                         extra  = interactions_long |> filter(status == "extra")  |> 
                           tidyr::pivot_wider(names_from = plant_taxa, values_from = n,
                                              values_fill = 0))

# remove unneccessary columns
grid_list_status_mat <- list(native = (grid_list_status[[1]][,-c(1:2)]),
                         intra = (grid_list_status[[2]][,-c(1:2)]),
                         extra = (grid_list_status[[3]][,-c(1:2)]))

# create interaction accumulation curve
sp_native <- specaccum(as.data.frame(grid_list_status_mat$native), method = "random")
sp_intra <- specaccum(as.data.frame(grid_list_status_mat$intra), method = "random")
sp_extra <- specaccum(as.data.frame(grid_list_status_mat$extra), method = "random")

# make it a data frame for plotting
accum_to_df <- function(x, status){data.frame(sites = x$sites,richness = x$richness,sd = x$sd,status = status )}
# apply function
df_accum <- bind_rows(accum_to_df(sp_native, "native"), accum_to_df(sp_intra,  "intra"),accum_to_df(sp_extra,  "extra"))
df_accum <- bind_rows(accum_to_df(sp_intra,  "intra"),accum_to_df(sp_extra,  "extra"))

# get accumulation curve
ACC_all <- ggplot(df_accum, aes(x = sites, y = richness, colour = status, fill = status)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = richness - sd,
                  ymax = richness + sd),
              alpha = 0.2,
              colour = NA) +
  ggpubr::theme_pubr() +
  labs(x = "Number of grid cells",
       y = "Accumulated unique interactions")+
  scale_color_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European"
                                               #, native = "Native"
                                               )) +
  scale_fill_manual(values = cols, labels = c(extra  = "Extra-European",intra  = "Intra-European"
                                              #, native = "Native"
                                              ))+
  theme(legend.position = "right")+
  scale_x_continuous(trans='log10') + annotation_logticks(sides = "b")

# save
ggsave("../Images/Accumulation_interactions_intra_extra.png", ACC_all, width= 6, height=3)
```


# 3 NETWORK
## 3.1 Metrics
```{r}
# check network characteristics
all <- networklevel(interaction_matrix[,-c(1:2)], index = c("connectance","NODF"))
all
# check by removing all alien species
nat <-  networklevel(interaction_matrix[interaction_matrix$status=="native",-c(1:2)], index = c("connectance","NODF"))
nat
# check by removing extra
nat_intra <-  networklevel(interaction_matrix[interaction_matrix$status!="extra",-c(1:2)], index = c("connectance","NODF"))
nat_intra
# check by removing intra
nat_extra <-  networklevel(interaction_matrix[interaction_matrix$status!="intra",-c(1:2)], index = c("connectance","NODF"))
nat_extra
```

## 3.2 Connectance
```{r}
# here calculate for removing all aliens

# create empty dataframe
data <- data.frame(nat= numeric(), all =numeric())
for(i in 1:1000){
  # select only the native
  nat <- interaction_matrix|> filter(status== "native") 
  # remove first two columns and the columns in which the colsum is 0
  nat <- nat[, -c(1:2)]
  nat <- nat[, colSums(nat)>0]
  # get dimension
  remove <- dim(nat)[1]
  # randomly sample exactly the same number of rows as were present before
  all <- interaction_matrix[sample.int(nrow(interaction_matrix), remove), -c(1:2)]
  # remove columns with colsum equal to 0
  all <- all[, colSums(all)>0]
  # get the minimum number of columns
  n_col <- min(ncol(nat), ncol(all))*0.9
  # sample both datasets again (number of plants is already equal, but number of bees varies)
  nat <- nat[, sample.int(ncol( nat), n_col)]
  all <- all[, sample.int(ncol( all), n_col)]
  # assign to the database
  data[i,] = data.frame(nat =  networklevel(nat, index ="connectance"), 
                        all = networklevel(all, index ="connectance" ))
}

# make data frame which facilites plotting and calculation of significance
data_long_all <- data |>  pivot_longer(cols = c(nat, all),
               names_to = "network",
               values_to = "connectance")
data_long_all$type <- "nat"


# plot
ggplot(data_long_all, aes(x = connectance, fill= network)) +
  geom_histogram(position = "identity", alpha= 1) +
  theme_pubr() +
  xlab("Connectance")+
  theme(legend.position = "right")


# check significance
test <- wilcox.test(connectance  ~ network, data_long_all)
if(test$p.value < 0.05) {
  means <- data_long_all |> group_by(network) |> summarise(mean = mean(connectance))|> 
    mutate(letter = ifelse(mean== max(mean), "a","b")) 
  data_long_all$letters <- means$letter[match(data_long_all$network, means$network)]
}

# significantly higher = more connected = higher number of interactions for dimension
```

```{r}
# here calculate for removing all aliens

# create empty dataframe
data_intra <- data.frame(nat= numeric(), all =numeric())
for(i in 1:1000){
  mat <- interaction_matrix|> filter(status!= "extra") 
  # select only the native
  nat <- mat|> filter(status== "native") 
  # remove first two columns and the columns in which the colsum is 0
  nat <- nat[, -c(1:2)]
  nat <- nat[, colSums(nat)>0]
  # get dimension
  remove <- dim(nat)[1]
  # randomly sample exactly the same number of rows as were present before
  all <- mat[sample.int(nrow(mat), remove), -c(1:2)]
  # remove columns with colsum equal to 0
  all <- all[, colSums(all)>0]
  # get the minimum number of columns
  n_col <- min(ncol(nat), ncol(all))*0.9
  # sample both datasets again (number of plants is already equal, but number of bees varies)
  nat <- nat[, sample.int(ncol( nat), n_col)]
  all <- all[, sample.int(ncol( all), n_col)]
  # assign to the database
  data_intra[i,] = data.frame(nat =  networklevel(nat, index ="connectance"), 
                        all = networklevel(all, index ="connectance" ))
}

# make data frame which facilites plotting and calculation of significance
data_long_intra <- data_intra |>  pivot_longer(cols = c(nat, all),
               names_to = "network",
               values_to = "connectance")
data_long_intra$type <- "nat_intra"

# plot
ggplot(data_long_intra, aes(x = connectance, fill= network)) +
  geom_histogram(position = "identity", alpha= 1) +
  theme_pubr() +
  xlab("Connectance")+
  theme(legend.position = "right")

# check significance
test <- wilcox.test(connectance  ~ network, data_long_intra)
if(test$p.value < 0.05) {
  means <- data_long_intra |> group_by(network) |> summarise(mean = mean(connectance))|> 
    mutate(letter = ifelse(mean== max(mean), "a","b")) 
  data_long_intra$letters <- means$letter[match(data_long_intra$network, means$network)]
}
# significantly higher = more connected = higher number of interactions for dimension
```

```{r}
# here calculate for removing all aliens

# create empty dataframe
data_extra <- data.frame(nat= numeric(), all =numeric())
for(i in 1:1000){
  mat <- interaction_matrix
  # select only the native
  nat <- mat|> filter(status!= "extra") 
  # remove first two columns and the columns in which the colsum is 0
  nat <- nat[, -c(1:2)]
  nat <- nat[, colSums(nat)>0]
  # get dimension
  remove <- dim(nat)[1]
  # randomly sample exactly the same number of rows as were present before
  all <- mat[sample.int(nrow(mat), remove), -c(1:2)]
  # remove columns with colsum equal to 0
  all <- all[, colSums(all)>0]
  # get the minimum number of columns
  n_col <- min(ncol(nat), ncol(all))*0.9
  # sample both datasets again (number of plants is already equal, but number of bees varies)
  nat <- nat[, sample.int(ncol( nat), n_col)]
  all <- all[, sample.int(ncol( all), n_col)]
  # assign to the database
  data_extra[i,] = data.frame(nat =  networklevel(nat, index ="connectance"), 
                        all = networklevel(all, index ="connectance" ))
}

# make data frame which facilites plotting and calculation of significance
data_long_extra <- data_extra |>  pivot_longer(cols = c(nat, all),
               names_to = "network",
               values_to = "connectance")
data_long_extra$type <- "nat_extra"

# plot
ggplot(data_long_extra, aes(x = connectance, fill= network)) +
  geom_histogram(position = "identity", alpha= 1) +
  theme_pubr() +
  xlab("Connectance")+
  theme(legend.position = "right")

# check significance
# check significance
test <- wilcox.test(connectance  ~ network, data_long_extra)
if(test$p.value < 0.05) {
  means <- data_long_all |> group_by(network) |> summarise(mean = mean(connectance))|> 
    mutate(letter = ifelse(mean== max(mean), "a","b")) 
  data_long_extra$letters <- means$letter[match(data_long_extra$network, means$network)]
}
# significantly higher = more connected = higher number of interactions for dimension
```



```{r}
data_long <- rbind(data_long_all, data_long_intra, data_long_extra)

# add spacers
spacers <- tibble( network   = NA,  connectance = NA, type = c("Spacer1", "Spacer2"))
plot_data <- bind_rows(data_long, spacers)

# add to order
order_levels <- c("nat_nat", "all_nat","NA_Spacer1" ,"nat_nat_extra","all_nat_extra",
                  "NA_Spacer2", "nat_nat_intra", "all_nat_intra") 
plot_data$combined <- paste(plot_data$network, plot_data$type, sep= "_")

# change order
plot_data$combined <- as.character(plot_data$combined)
# set reversed order
plot_data$combined <- factor(plot_data$combined,levels = rev(order_levels))
plot_data$names <- plot_data$combined

change_labels <-c("nat_nat"= "Only natives",
                  "all_nat"= "All species",
                  "NA_Spacer1"="" ,
                  "nat_nat_extra"= "Natives and\nIntra-European aliens",
                  "all_nat_extra"= "All species",
                  "NA_Spacer2"="", 
                  "nat_nat_intra"= "Only natives", 
                  "all_nat_intra"="Natives and\nIntra-European aliens")

levels(plot_data$names) <- change_labels[levels(plot_data$names)]
plot_labels <- plot_data[!duplicated(plot_data$combined),4:5]

cols_network <- c(
  "Only natives"                        = "#0072B2",  # deep blue
  "Natives and\nIntra-European aliens"   = "#CC79A7",  # magenta
  "All species"                         = "#E69F00"   # amber
)

# plot data
plot_con <- ggplot(plot_data, aes(x= connectance, y= combined, fill= names))+
  stat_density_ridges(alpha = 0.8, rel_min_height = 0.01, scale = 1.2,
                      quantile_lines = TRUE, quantile_fun = mean, na.rm = TRUE)+
  theme_ridges() +
  theme_pubr() +
  scale_y_discrete(labels = change_labels)+
  geom_text(data = plot_labels, aes(x = max(plot_data$connectance, na.rm = TRUE), 
                                    y = combined, label = letters),inherit.aes = FALSE,
             hjust = 0,size = 4 )+
  theme(legend.position = "none", axis.title.y = element_blank(), axis.title.x = element_text(size = 14),
        axis.text.y  = element_text(size = 12), axis.text.x  = element_text(size = 12), axis.ticks.y = element_blank() ) +
  xlab("Connectance")+
  scale_y_discrete(labels = change_labels)+
  scale_fill_manual(values = cols_network)

ggsave("../Images/Connectance_overall_status.png", plot_con, width =6, height=4)
```

## 3.3 NODF
```{r}
# here calculate for removing all aliens
i=1
# create empty dataframe
data <- data.frame(nat= numeric(), all =numeric())
for(i in 1:50){
  # select only the native
  nat <- interaction_matrix|> filter(status== "native") 
  # remove first two columns and the columns in which the colsum is 0
  nat <- nat[, -c(1:2)]
  nat <- nat[, colSums(nat)>0]
  # get dimension
  remove <- dim(nat)[1]
  # randomly sample exactly the same number of rows as were present before
  all <- interaction_matrix[sample.int(nrow(interaction_matrix), remove), -c(1:2)]
  # remove columns with colsum equal to 0
  all <- all[, colSums(all)>0]
  # get the minimum number of columns
  n_col <- min(ncol(nat), ncol(all))*0.9
  # sample both datasets again (number of plants is already equal, but number of bees varies)
  nat <- nat[, sample.int(ncol( nat), n_col)]
  all <- all[, sample.int(ncol( all), n_col)]
  # assign to the database
  data[i,] = data.frame(nat =  networklevel(nat, index ="NODF"), 
                        all = networklevel(all, index ="NODF" ))
}

table(interaction_matrix$status)

# make data frame which facilites plotting and calculation of significance
data_long_all <- data |>  pivot_longer(cols = c(nat, all),
               names_to = "network",
               values_to = "NODF")
data_long_all$type <- "nat"


# plot
ggplot(data_long_all, aes(x = NODF, fill= network)) +
  geom_histogram(position = "identity", alpha= 1) +
  theme_pubr() +
  xlab("NODF")+
  theme(legend.position = "right")


# check significance
test <- wilcox.test(NODF  ~ network, data_long_all)
if(test$p.value < 0.05) {
  means <- data_long_all |> group_by(network) |> summarise(mean = mean(NODF))|> 
    mutate(letter = ifelse(mean== max(mean), "a","b")) 
  data_long_all$letters <- means$letter[match(data_long_all$network, means$network)]
}

# significantly higher = more connected = higher number of interactions for dimension
```

```{r}
# here calculate for removing all aliens

# create empty dataframe
data_intra <- data.frame(nat= numeric(), all =numeric())
for(i in 16:50){
  mat <- interaction_matrix|> filter(status!= "extra") 
  # select only the native
  nat <- mat|> filter(status== "native") 
  # remove first two columns and the columns in which the colsum is 0
  nat <- nat[, -c(1:2)]
  nat <- nat[, colSums(nat)>0]
  # get dimension
  remove <- dim(nat)[1]
  # randomly sample exactly the same number of rows as were present before
  all <- mat[sample.int(nrow(mat), remove), -c(1:2)]
  # remove columns with colsum equal to 0
  all <- all[, colSums(all)>0]
  # get the minimum number of columns
  n_col <- min(ncol(nat), ncol(all))*0.9
  # sample both datasets again (number of plants is already equal, but number of bees varies)
  nat <- nat[, sample.int(ncol( nat), n_col)]
  all <- all[, sample.int(ncol( all), n_col)]
  # assign to the database
  data_intra[i,] = data.frame(nat =  networklevel(nat, index ="NODF"), 
                        all = networklevel(all, index ="NODF" ))
}

# make data frame which facilites plotting and calculation of significance
data_long_intra <- data_intra |>  pivot_longer(cols = c(nat, all),
               names_to = "network",
               values_to = "NODF")
data_long_intra$type <- "nat_intra"

# plot
ggplot(data_long_intra, aes(x = NODF, fill= network)) +
  geom_histogram(position = "identity", alpha= 1) +
  theme_pubr() +
  xlab("NODF")+
  theme(legend.position = "right")

# check significance
test <- wilcox.test(NODF  ~ network, data_long_intra)
if(test$p.value < 0.05) {
  means <- data_long_intra |> group_by(network) |> summarise(mean = mean(NODF))|> 
    mutate(letter = ifelse(mean== max(mean), "a","b")) 
  data_long_intra$letters <- means$letter[match(data_long_intra$network, means$network)]
}
# significantly higher = more connected = higher number of interactions for dimension
```

```{r}
# here calculate for removing all aliens

# create empty dataframe
data_extra <- data.frame(nat= numeric(), all =numeric())
for(i in 1:50){
  mat <- interaction_matrix
  # select only the native
  nat <- mat|> filter(status!= "extra") 
  # remove first two columns and the columns in which the colsum is 0
  nat <- nat[, -c(1:2)]
  nat <- nat[, colSums(nat)>0]
  # get dimension
  remove <- dim(nat)[1]
  # randomly sample exactly the same number of rows as were present before
  all <- mat[sample.int(nrow(mat), remove), -c(1:2)]
  # remove columns with colsum equal to 0
  all <- all[, colSums(all)>0]
  # get the minimum number of columns
  n_col <- min(ncol(nat), ncol(all))*0.9
  # sample both datasets again (number of plants is already equal, but number of bees varies)
  nat <- nat[, sample.int(ncol( nat), n_col)]
  all <- all[, sample.int(ncol( all), n_col)]
  # assign to the database
  data_extra[i,] = data.frame(nat =  networklevel(nat, index ="NODF"), 
                        all = networklevel(all, index ="NODF" ))
}

# make data frame which facilites plotting and calculation of significance
data_long_extra <- data_extra |>  pivot_longer(cols = c(nat, all),
               names_to = "network",
               values_to = "NODF")
data_long_extra$type <- "nat_extra"

# plot
ggplot(data_long_extra, aes(x = NODF, fill= network)) +
  geom_histogram(position = "identity", alpha= 1) +
  theme_pubr() +
  xlab("NODF")+
  theme(legend.position = "right")

# check significance
# check significance
test <- wilcox.test(NODF  ~ network, data_long_extra)
if(test$p.value < 0.05) {
  means <- data_long_all |> group_by(network) |> summarise(mean = mean(NODF))|> 
    mutate(letter = ifelse(mean== max(mean), "a","b")) 
  data_long_extra$letters <- means$letter[match(data_long_extra$network, means$network)]
}
# significantly higher = more connected = higher number of interactions for dimension
```



```{r}
data_long <- rbind(data_long_all, data_long_intra, data_long_extra)

# add spacers
spacers <- tibble( network   = NA,  NODF = NA, type = c("Spacer1", "Spacer2"))
plot_data <- bind_rows(data_long, spacers)

# add to order
order_levels <- c("nat_nat", "all_nat","NA_Spacer1" ,"nat_nat_extra","all_nat_extra",
                  "NA_Spacer2", "nat_nat_intra", "all_nat_intra") 
plot_data$combined <- paste(plot_data$network, plot_data$type, sep= "_")

# change order
plot_data$combined <- as.character(plot_data$combined)
# set reversed order
plot_data$combined <- factor(plot_data$combined,levels = rev(order_levels))
plot_data$names <- plot_data$combined

change_labels <-c("nat_nat"= "Only natives",
                  "all_nat"= "All species",
                  "NA_Spacer1"="" ,
                  "nat_nat_extra"= "Natives and\nIntra-European aliens",
                  "all_nat_extra"= "All species",
                  "NA_Spacer2"="", 
                  "nat_nat_intra"= "Only natives", 
                  "all_nat_intra"="Natives and\nIntra-European aliens")

levels(plot_data$names) <- change_labels[levels(plot_data$names)]
plot_labels <- plot_data[!duplicated(plot_data$combined),4:5]

cols_network <- c(
  "Only natives"                        = "#0072B2",  # deep blue
  "Natives and\nIntra-European aliens"   = "#CC79A7",  # magenta
  "All species"                         = "#E69F00"   # amber
)

# plot data
plot_NODF <- ggplot(plot_data, aes(x= NODF, y= combined, fill= names))+
  stat_density_ridges(alpha = 0.8, rel_min_height = 0.01, scale = 1.2,
                      quantile_lines = TRUE, quantile_fun = mean, na.rm = TRUE)+
  theme_ridges() +
  theme_pubr() +
  scale_y_discrete(labels = change_labels)+
  geom_text(data = plot_labels, aes(x = max(plot_data$NODF, na.rm = TRUE)*1.05, 
                                    y = combined, label = letters),inherit.aes = FALSE,
             hjust = 0,size = 4 )+
  theme(legend.position = "none", axis.title.y = element_blank(), axis.title.x = element_text(size = 14),
        axis.text.y  = element_text(size = 12), axis.text.x  = element_text(size = 12), axis.ticks.y = element_blank() ) +
  xlab("NODF")+
  scale_y_discrete(labels = change_labels)+
  scale_fill_manual(values = cols_network)

ggsave("../Images/NODF_overall_status.png", plot_NODF, width =6, height=4)
```


## 3.4 Modularity
```{r}
# check contribution species
# some intra come multiple times (intra and native)
# merge status and plant name if this is the case
dup <- interactions |> distinct(plant_taxa, status)
dup <- dup[duplicated(dup$plant_taxa) | duplicated(dup$plant_taxa, fromLast = T),]

dup$new_name <- paste(dup$plant_taxa, " (", dup$status,")", sep="")

interactions$plant_new_name <-  paste(interactions$plant_taxa, " (", interactions$status,")",
                                      sep="")

# create matrix to use
mat <- interactions |>
  # get only unique interactions
  distinct(plant_new_name, name) |> 
  # give these interactions all a 1 (presence)
  mutate(value = 1) |>
  # assign names
  pivot_wider(names_from  = name,  values_from = value, values_fill = 0) |>
  column_to_rownames("plant_new_name") |>
  as.matrix()

# check whether dimensions are okay
any(duplicated(rownames(mat)))
any(is.na(mat))
```


```{r}
# run modules assessment
res <- netcarto(mat, bipartite=T)
# save (takes very long to calculate)
write_rds(res, "../Extra/Intermediate/modularity_overall.Rdata")

# retrieve data
species_mod <- res[[1]]
# change again status
species_mod <- species_mod |> tidyr::extract(name, into = c("plant_taxa", "status"), regex = "^(.*) \\((.*)\\)$" )

# check boundaries
quantile(species_mod$connectivity, 0.9)
quantile(species_mod$participation, 0.9)
```


```{r}
# for plotting, we will change the values
species_mod$alpha_val <- ifelse(species_mod$status == "native", 0.75, 1)
species_mod$size_val <- ifelse(species_mod$status == "native", 1,2)

# plot
p <- ggplot(species_mod, aes(participation, connectivity, col=status))+
  geom_point(alpha= species_mod$alpha_val, size= species_mod$size_val)+
  theme_pubr()+
  geom_vline(xintercept=quantile(species_mod$participation, 0.9), linetype='dotted', alpha=1)+
  geom_hline(yintercept=quantile(species_mod$connectivity, 0.9), linetype='dotted', alpha=1)+
  xlab("Among-module connectivity (c)")+
  ylab("Within-module degree (z)")+
  scale_colour_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native"), 
                    labels = c("Extra-\nEuropean", "Intra-\nEuropean", "Native")) +
  scale_fill_manual(values=c( "#F07167","#7CAE00", "#00A9A5"), 
                    breaks=c("extra","intra", "native"), 
                    labels = c("Extra-\nEuropean", "Intra-\nEuropean", "Native"))+
  theme(legend.position = "bottom")+ 
  theme(legend.title = element_blank())

save <- ggExtra::ggMarginal(p, type = "boxplot", groupColour = TRUE, groupFill = TRUE, margins = "both")

# save
ggsave("../Images/Modularity_overall.png", save, width = 6, height = 4)

# check significance participation
sign_status <- (lm(participation ~ status, species_mod))# check significance
emm <- emmeans(sign_status, pairwise ~ status)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
cld_results

# check significance connectoivity 
sign_status <- (lm(connectivity ~ status, species_mod))# check significance
emm <- emmeans(sign_status, pairwise ~ status)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
cld_results

# in future add this manually
```


```{r}
# look at modules
length(unique(species_mod$module))

# look at modules
check_rel <- species_mod |> group_by(module) |> summarise(rel_nat=sum(status=="native")/length(status), n=n())

ggplot(species_mod)+
  geom_violin(aes( y=n))

check_rel <- check_rel[check_rel$n > 1,]

# check this function
computeModules()
```



# 4 Plot
```{r}
# create matrix 
plot_matrix <- interactions |> group_by(status, pol_family, plant_taxa, name) |> 
  summarise(n =n())
# count number of unique interactions
plot_matrix <- interactions |> group_by(status, pol_family) |> 
  summarise(n =n())

# make dataframe
plot_matrix <- plot_matrix |> pivot_wider(values_from = n, names_from = pol_family) |>
  as.data.frame()
# change names and remove status column
rownames(plot_matrix) <- c("Extra-European","Intra-European","Native")
plot_matrix <- plot_matrix[, -1]

# assign colour
bee_cols <- c("Andrenidae"= "#E69F00", "Apidae"= "#0072B2", "Halictidae" = "#CC79A7","Megachilidae"= "#D55E00","Melittidae"= "#6A3D9A" )
cols <- c("Extra-European" = "#F07167","Intra-European"  = "#7CAE00","Native" = "#00A9A5")

# plot
png("../Images/my_plotweb.png", width = 1500, height = 750, res = 150)
plotweb(plot_matrix, lower_color = cols, link_color = "lower", higher_color = bee_cols, text_size = 0.85)
dev.off()
```


# ECO
# 1 Subdivide
```{r}
# remove singletons
interactions <- interactions |>   group_by(name, plant_taxa, eco_region) |>  
  filter(n() >= 2) |>  ungroup()

# only status subdivisions
interaction_matrix <-interactions|>
  dplyr::count(name,plant_taxa,eco_region,status)|> 
  tidyr::pivot_wider(names_from = name,values_from = n, values_fill = list(n = 0))

interaction_matrix <- as.data.frame(interaction_matrix)
```


```{r}
# create databases per network
ecoregions <- interactions |> group_by(eco_region) |> summarise(n=n() )
ecoregions_int <- interaction_matrix |> group_by(eco_region) |>
  summarise(n_int=n(), n_alien = sum(status != "native"), n_nat = sum(status == "native"))
ecoregions <- left_join(ecoregions, ecoregions_int)

# how well are the two correlated
cor.test(ecoregions$n, ecoregions$n_int, method= "pearson")

# for which will we calculate the network metrics
boxplot(ecoregions$n_int)
summary(ecoregions$n_int)

# let's say for now a minimum of 10 interactions
min_int <- 100
```


```{r}
# get abbreviations for the names in the database
eco <- data.frame(unique(ecoregions$eco_region))
colnames(eco) <- "name"
eco$abbreviation <- c("Turk", "Alps", "Balt","Bor","Celt","Centr","Engl","Mix","Med",
                            "Spain","Pan","Sam","Ib","Broad","Con","Grass","Tun","West") 
ecoregions$abbreviation <- eco$abbreviation[match(ecoregions$eco_region, eco$name)]

# remove some observations from this list which are too small to allow calculation
ecoregions <- ecoregions[ecoregions$n_int >= min_int,]

# add abbreviations
interaction_matrix$eco_region <- ecoregions$abbreviation[match(interaction_matrix$eco_region,
                                                               ecoregions$eco_region)]

# remove interactions not present anymore
interaction_matrix <- interaction_matrix[!is.na(interaction_matrix$eco_region),]

# make the name column better to discern between intra and native
interaction_matrix$name <-  paste(interaction_matrix$plant_taxa, 
                                            " (", interaction_matrix$status,")", sep="")
interaction_matrix <- interaction_matrix |> relocate(name, .after= status)
```


```{r}
# subdivide interaction matrices
interaction_matrices <- list()
for(i in ecoregions$abbreviation){
  p <- (interaction_matrix[interaction_matrix$eco_region == i, , drop = FALSE])
  rownames(p) <- p$name
  p <- p[, 5:ncol(p)]
  p <- p[, colSums(p) > 0]
  interaction_matrices[[i]] <- p
}
```


# 2 Species
## 2.1 Metrics
```{r}
# get species level networks
species_degree_df <- lapply(names(interaction_matrices), function(i){

  mat <- interaction_matrices[[i]]
  mat <- as.matrix(mat)
  mode(mat) <- "numeric"
  nL <- nrow(mat)
  nH <- ncol(mat)

  deg <- specieslevel(mat,"degree")
  ND <- specieslevel(mat,"normalised degree")
  strg <- specieslevel(mat,"species strength")

  if(nL>=2 && nH>=2){
    NSI <- specieslevel(mat,"NSI")
    BC <- specieslevel(mat,"betweenness")
    NSI_H <- NSI$`higher level`
    NSI_L <- NSI$`lower level`
    BC_H <- BC$`higher level`$betweenness
    BC_L <- BC$`lower level`$betweenness
  } else {
    NSI_H <- rep(NA,nH)
    NSI_L <- rep(NA,nL)
    BC_H <- rep(NA,nH)
    BC_L <- rep(NA,nL)
  }
  
  bind_rows(data.frame(species=colnames(mat), n=nH, degree=deg$`higher level`,
                       ND=ND$`higher level`, strength=strg$`higher level`, NSI=NSI_H, 
                       BC=BC_H, trophic_level="pollinator", ecoregion=i),
            data.frame(species=rownames(mat), n=nL, degree=deg$`lower level`,
                       ND=ND$`lower level`, strength=strg$`lower level`, NSI=NSI_L, 
                       BC=BC_L, trophic_level="plant", ecoregion=i))
}) |> bind_rows()

# make nicer row names
rownames(species_degree_df) <- 1: nrow(species_degree_df)
```



# ECO-LAND
# 1 Subdivide
```{r}
# remove singletons
interactions <- interactions |>   group_by(name, plant_taxa, eco_region, euromed_region) |>  
  filter(n() >= 2) |>  ungroup()

# only status subdivisions
interaction_matrix <-interactions|>
  dplyr::count(name,plant_taxa,eco_region,euromed_region,status)|> 
  tidyr::pivot_wider(names_from = name,values_from = n, values_fill = list(n = 0))

interaction_matrix <- as.data.frame(interaction_matrix)
```


```{r}
# create databases per network
ecoregions <- interactions |> group_by(eco_region, euromed_region) |> summarise(n=n() )
ecoregions_int <- interaction_matrix |> group_by(eco_region, euromed_region) |>
  summarise(n_int=n(), n_alien = sum(status != "native"), n_nat = sum(status == "native"))
ecoregions <- left_join(ecoregions, ecoregions_int)

# how well are the two correlated
cor.test(ecoregions$n, ecoregions$n_int, method= "pearson")

# for which will we calculate the network metrics
boxplot(ecoregions$n_int)
summary(ecoregions$n_int)

# let's say for now a minimum of 10 interactions
min_int <- 10
```


```{r}
# get abbreviations for the names in the database
eco <- data.frame(unique(ecoregions$eco_region))
colnames(eco) <- "name"
eco$abbreviation <- c("Turk", "Alps", "Balt","Bor","Celt","Centr","Engl","Mix","Med",
                            "Spain","Pan","Sam","Ib","Broad","Con","Grass","Tun","West") 
ecoregions$abbreviation <- eco$abbreviation[match(ecoregions$eco_region, eco$name)]
ecoregions$name <- paste(ecoregions$abbreviation, ecoregions$euromed_region, sep= "_")

# remove some observations from this list which are too small to allow calculation
ecoregions <- ecoregions[ecoregions$n_int >= 10,]

# add abbreviations
interaction_matrix$eco_region <- ecoregions$abbreviation[match(interaction_matrix$eco_region,
                                                               ecoregions$eco_region)]

# link with region
interaction_matrix$region <- paste(interaction_matrix$eco_region,
                                   interaction_matrix$euromed_region, sep= "_")
interaction_matrix <- interaction_matrix |> relocate(region, .after= euromed_region)
```


```{r}
# subdivide interaction matrices
interaction_matrices <- list()
for(i in ecoregions$abbreviation){
  p <- interaction_matrix[interaction_matrix$eco_region == i, , drop = FALSE]
  rownames(p) <- p$plant_taxa
  p <- p[, -c(1:2), drop = FALSE]
  p <- p[, colSums(p) > 0, drop = FALSE]
  interaction_matrices[[i]] <- p
}
```


# 2 Species
## 2.1 Metrics
```{r}
# get species level networks
species_degree_df <- lapply(names(interaction_matrices), function(i){

  mat <- interaction_matrices[[i]]
  mat <- as.matrix(mat)
  mode(mat) <- "numeric"
  nL <- nrow(mat)
  nH <- ncol(mat)

  deg <- specieslevel(mat,"degree")
  ND <- specieslevel(mat,"normalised degree")
  strg <- specieslevel(mat,"species strength")

  if(nL>=2 && nH>=2){
    NSI <- specieslevel(mat,"NSI")
    BC <- specieslevel(mat,"betweenness")
    NSI_H <- NSI$`higher level`
    NSI_L <- NSI$`lower level`
    BC_H <- BC$`higher level`$betweenness
    BC_L <- BC$`lower level`$betweenness
    n_contr <- nestedcontribution(mat, nsimul = 99)
    n_contr_L <- n_contr$`lower level`
    n_contr_H <- n_contr$`higher level`
  } else {
    NSI_H <- rep(NA,nH)
    NSI_L <- rep(NA,nL)
    BC_H <- rep(NA,nH)
    BC_L <- rep(NA,nL)
    n_contr_L <- rep(NA,nH)
    n_contr_H <- rep(NA,nH)
  }
  
  bind_rows(
    data.frame(species=colnames(mat), n=nH,degree=deg$`higher level`,ND=ND$`higher level`, strength=strg$`higher level`, NSI=NSI_H, BC=BC_H, 
               n_contr_H = n_contr_H,trophic_level="pollinator", ecoregion=i),
    data.frame(species=rownames(mat), n=nL, degree=deg$`lower level`,ND=ND$`lower level`, strength=strg$`lower level`, NSI=NSI_L, BC=BC_L,
               n_contr_L = n_contr_L, trophic_level="plant", ecoregion=i))
}) |> bind_rows()

rownames(species_degree_df) <- 1: nrow(species_degree_df)
species_degree_df <- species_degree_df[, -10]
species_degree_df$node.specialisation.index.NSI[species_degree_df$node.specialisation.index.NSI=="NaN"] <- NA
```


```{r}
# assign family and genus levels
species_degree_df$genus <- interactions$plant_genus[match(species_degree_df$species, interactions$plant_taxa)]
species_degree_df$family <- interactions$plant_family[match(species_degree_df$species, interactions$plant_taxa)]
species_degree_df$genus[is.na(species_degree_df$genus)] <- interactions$pol_genus[match(species_degree_df$species[is.na(species_degree_df$genus)], interactions$name)]
species_degree_df$family[is.na(species_degree_df$family)] <- interactions$pol_family[match(species_degree_df$species[is.na(species_degree_df$family)], interactions$name)]
```


## 2.2 Plot
```{r}
# we will remove the datasets with too little data
species_degree_df <- species_degree_df[species_degree_df$n > 10,]

# plot plant and pollinator degrees
ggplot(species_degree_df, aes(x=trophic_level, y=PSI, fill=trophic_level)) + 
    geom_boxplot() +
    facet_wrap(~ecoregion, scale="free")

# plot normalised degree for plants
ggplot(species_degree_df[species_degree_df$trophic_level== "plant",], aes(x=node.specialisation.index.NSI, y= ecoregion, fill = ecoregion)) + 
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")
```


## 2.3 Significance
```{r}
# are the differences among groups significant
sign_ecoregion_pol <- glmer(node.specialisation.index.NSI ~ ecoregion + (1|family/genus ), species_degree_df[species_degree_df$trophic_level=="plant",], family = gaussian(link = "identity"))
summary(sign_ecoregion_pol)
plot(simulateResiduals(sign_ecoregion_pol))

# check significance
emm <- emmeans(sign_ecoregion_pol, pairwise ~ecoregion)
cld_results <- cld(emm, alpha = 0.05, Letters = letters, adjust = "tukey", reversed=T)
cld_results$.group <- gsub(" ","",cld_results$.group)
cld_results
```

# 3 Network
```{r}

```


# 4 Module
```{r}
mod <- computeModules(interaction_matrices[[2]])
plotModuleWeb(mod)
```




# FIELDWORK
```{r}
# check some species
check <- c( "Allium sphaerocephalon", "Allium ursinum", "Allium vineale", "Arnica montana", "Campanula glomerata","Campanula persicifolia",
            "Campanula rotundifolia", "Campanula trachelium", "Centaurea jacea","Centaurea scabiosa",  "Cichorium intybus", "Cirsium vulgare",
            "Echium vulgare",  "Eupatorium cannabinum", "Helichrysum arenarium","Inula salicina", "Jacobaea vulgaris","Knautia arvensis",
            "Lotus corniculatus", "Origanum vulgare", "Picris hieracioides","Reseda lutea", "Scabiosa columbaria","Scorzonera humilis",
            "Serratula tinctoria","Tanacetum vulgare",  "Tragopogon pratensis","Trifolium pratense", "Succisa pratensis","Potentilla erecta",
            "Lythrum salicaria","Odontites vulgaris","Lysimachia vulgaris")

subset <- interactions[interactions$plant_taxa %in% check  & interactions$euromed_region=="Belgium",]
subset <- subset |> distinct(plant_taxa, name) |> group_by(plant_taxa) |> summarise(n=n())

# check again
pieter <- readxl::read_excel("C:/Users/u0166342/Documents/Doctoraat/Camera/Species/Species_Pieter_Vannormelingen.xlsx", sheet= "DirkM_14apr2025")

subset <- interactions[interactions$name %in% pieter$`Scientific name`[pieter$flower_pref=="oligo"] &
                       interactions$plant_taxa %in% check & interactions$euromed_region=="Belgium", ]
subset <- subset |> distinct(plant_taxa, name) |> group_by(plant_taxa) |> summarise(n=n())
```


